### 2017-05-08 manuscript.Rmd
Residual commit

---
title: "Renvisioning a future in scholarly communication"
author: "CHJ Hartgerink"
date: "`r Sys.Date()`"
output: html_document
toc: true
fig_caption: true
bibliography: library.bib
---

__For the 2017 IFLA conference, theme: The Open Revolution__

## Abstract

## Body

Scholarly communication is in dire need of disruption. People often tell me we need incremental change from within to produce a sustainable, modern scholarly communication system, but I oppose this idea. After decades of identifying various problems we have seen little sustainable change that truly and deeply reimagines how we communicate. Yes, PLOS changed the game. Yes, RIO journal is one of the most innovative ideas in scholarly communication of the last two years. But we are still trying to turn a horse (i.e., paper based communication) into a motor vehicle (i.e., the new thing).

In this perspective piece, I try to radically reimagine the scholarly communications infrastructure based on the tools that are available to us now.  I want these ideas to be criticized; only with diversity of ideas can we truly make progress. Currently the decisions are driven by a few oligopolies [@10.1371/journal.pone.0127502] or the tenured professors, the policymakers, and other people who are actually farthest removed from doing actual research. Many of these people stem from a different generation and have become successful in the broken system we are now stuck with. As a result, they might underestimate some of the problems we now face. This inherently decreases the diversity of ideas and stalls the development of how scholars communicate, with dire consequences (as we noticed in recent years, with unfounded and unconstructive criticisms of replication initiatives).

In the last decades, interpersonal communication has changed tremendously. These past decades have brought with it much diversity in communication, such that it can become confusing (see Figure 1). At the same time, the diversity is liberating by allowing innovative ways to communicate (e.g., social media) or making available previously difficult to use safety measures (e.g., encryption). Diversity has thus spawned new ways of communicating or have made previously esoteric communication tools available to the masses.

```{r xkcd-1810, fig.cap="_Figure 1._ A comical example of the diversity in communication channels.", out.width="60%", fig.align="center", echo = FALSE}
knitr::include_graphics('figs/xkcd_1810-cc-by-nc-2_5.png', auto_pdf = TRUE)
```

However, the way scholars communicate their findings has barely changed; we are still stuck in a paper paradigm, albeit writing on digital pages. In recent years, options have started to arise with tools such as [Rmarkdown]() and [Jupyter notebooks]() to break free from this paper paradigm. However, in the end, the scholarly communications system as we now know it still requires submitting to a journal that fits all of it back into a page-by-page, static, and paper format. This not only prevents new ways of communicating findings, it also prevents encouragement of discussion amongst readers about the findings, but to passively consume. Not to even start of the antiquated and wholly redundant process of waiting for the article to be assigned an issue number.

The scholarly process is much more complicated than can be captured in a static and retrospective publication. For example, a quantitative scholar needs to develop methods to investigate a research question. Feedback from peers would be valuable at that point. This has slowly been picked up with so called [Registered Reports](). However, this is more the exception than the norm. Additionally, researchers create materials, protocols, and other outputs during the research process that now are not (aptly) sharable in a static publication.

The scholarly process has also grown even _more_ complicated due to increased variation in the research tools available. Point and click software has increased the number of statistical analyses a researcher can do within minutes; the development of technology has created new types of data output that greatly exceed human comprehension and require algorithms before being able to make sense of the data (e.g., fMRI data). How a researcher applies all these tools is essential, but the current scholarly communication system does not require the details. Lack of direct reproducibility of results [e.g., @10.1101/056473] therefore is no surprise when we do not know how data are processed and analyzed.

As such, the scholarly process has become _more_ of a black box than it was before. Increasing complexity and lack of transparency makes it very difficult to reproduce steps. It is similar to being dropped in the middle of a desert with a blindfold on, and then being asked to find your way back home. Moreover, much of the software is proprietary and it is simply impossible to verify results or understand how they originated [@10.1038/nature10836]. Trying to reverse engineer these pieces of software is actually often illegal based on the unilateral agreement that you have to accept before using the software. As such, it is practically impossible to understand how software development affects how results are computed, while testing for this is easy with open-source software.
<!-- Dit moet nog herschreven worden -->

Due to these problems, breaking up the reign of the paper paradigm is a necessity to better represent and understand how scholarly results come into existence. Of course paper is a valid medium and should not be neglected, but this is not the case when scholarly communications becomes truly digitally born. PDF versions of outputs can still be produced.

Additionally, the way we organize scholarly communication cannot break free from the legacies if we do not radically reenvision it. due to the commodification of the resource (i.e., knowledge), a new system where ways to consume information is commodified are actively stifled.

Commodification of consumption greatly encourages society to innovate how people can discover and reuse information. <!-- For Wikipedia alone, one of the greatest resources of uncommodified knowledge, there are tons of applications that allow you to consume that information in various ways on various devices.  -->

By focusing on the commodification of consumption instead of commodification of the resource, the problem of access to knowledge can be resolved in a sustainable manner. <!-- Politicians and scholars have been discussing Open Access for at least 15 years. -->

Additionally, commodification of consumption removes several perverse incentives from the scholarly system that now produces unreliable knowledge. <!-- For example, innovativeness and surprisingness are criteria upon which articles are currently selected, where those exact some properties increase the probability that the finding is false [@10.1371/journal.pmed.0020124]. -->

The main tenet underlying the themes of Open Access, Open Data, Open Science, and replication initiatives in scholarly communication is sustainability. <!-- Sustainability of consuming knowledge (Open Access), sustainability of producing reliable knowledge ("reproducibility crisis"), sustainability of The serial crisis indicates that access to knowledge is currently unsustainable (e.g., Harvard cannot even access everything with its massive budget).  -->

In this perspective piece, I propose a redesign of the scholarly communication system that is flexible, decentralized, and distributed, which immediately allows for more sustainability and diversity in the way we consume, produce, and access knowledge.

## Sustainable scholarly process

Single points of failures make a system vulnerable to outages, (malicious) adjustment of content, or even complete removal of content. <!-- As such, a centralized infrastructure is easier to disrupt than a decentralized or distributed one. The Internet is currently organized in such a way that it is relatively centralized, whereas the Darknet (i.e., an alternative cyberspace) is less centralized and has been found to be more resilient [@10.1103/PhysRevE.95.022313].  -->

The scholarly process is ridden with single or few points of failures and sharing makes this a many points for failure system. <!-- For exactly this reason it is so important that scholars become more transparent about their data, their hypotheses, their results, and the chronology of a. -->

<!-- The scholarly communications system is currently centralized at publisher websites, is insecure, and is not resilient. HTTP(S) because redistribution is frequently by copyright Unrestrictive licensing, such as the Creative Commons BY license, allow for free redistribution and increases persistence. For example, anyone can download the entire corpus of Hindawi right now and if Hindawi would then become inaccessible, the articles would still be available regardless. Copyright creates dependence whereas the lack thereof creates independence. -->

Hence, a sustainable scholarly communications system needs to be both decentralized and distributed.

Decentralization is achievable with new peer-to-peer (p2p) Internet protocols such as `dat` and `ipfs`. <!-- Such p2p networks send information across the network, but are is resilient to nodes being removed because they operate in mesh network. If 20 peers have the files, removing one from the network does not affect the availability of the files-->

Such p2p networks also allow for ready redistribution when copyright is leveraged to work for distribution.

Alongside decentralization and distribution, integrity of the scholarly record can also become better. The chronology of the scholarly process can be ensured by splitting up the paper article as we know it into nodes of information.

These nodes can easily be logged on a blockchain based ledger of which everyone can have a copy (e.g., [Hyperledger](https://github.com/hyperledger/fabric)-based). <!-- An example is available below -->

```json
{
	"parent": null,
	"self": "dat://92b9b1b95e18c1dd1546c25ac49609da6fb0e3f2abc94ecf17609c565888c16b",
	"time-deposit": "2017-04-10T18:38:03+02:00",
	"time-available": "2017-04-10T18:38:03+02:00",
	"authors": "b601f276870ebf4920b3c78c4af51d3ed3bf4db7ab2b9c4faa74e0b457aa13ed",
	"type": "hypotheses"
}

{
	"parent": "dat://92b9b1b95e18c1dd1546c25ac49609da6fb0e3f2abc94ecf17609c565888c16b",
	"self": "dat://c2c291f93fa635c7fc7bc6e8ba504937b1821c5be45ab7c859f6e6cd19a291b7",
	"time-deposit": "2017-04-12T12:07:06+02:00",
	"time-available": "2017-08-12T12:07:06+02:00",
	"authors": "b601f276870ebf4920b3c78c4af51d3ed3bf4db7ab2b9c4faa74e0b457aa13ed",
	"type": "data"
}
```

Each entry in the ledger needs to be self-containing and refer to all relevant files.

A parent node can never occur after a child node,

The ledger can also allow for time-locked sharing but with a clear chronology of creation.

Nodes cannot be made available

Combinations of ORCIDs can be hashed into unique author combinations (and ordering!) to provide identification over time, but remove gender and status bias (amongst other things).

What types of nodes are contained in the ledger, is something that requires discussion and can change over time. Nonetheless, the fact remains that if someone presents results without

For a sustainable distribution system, persistent nodes need to be identified: libraries.

Nonetheless, other stakeholders benefiting from maintaining the distribution of the decentralized network are (not-)for-profit companies that build services on top of the resources.

## Locking the system open

When Brewster Kahle spoke at OpenCon2016, he mentioned that we need to lock the system open. At first, it would seem only the open advocates win in that case, but I think the for-profit companies can benefit from this as well by opening up a whole new market where more return can be made and in a more sustainable manner.

## References

### 2017-05-10 manuscript.Rmd
Draft first section paper

@@
 ---
-title: "Renvisioning a future in scholarly communication"
-author: "CHJ Hartgerink"
-date: "`r Sys.Date()`"
-output: html_document
-toc: true
-fig_caption: true
+title: "Re-envisioning a future in scholarly communication"
+author: "Chris H.J. Hartgerink"
+date: '`r Sys.Date()`'
+output:
+  html_document:
+    includes:
+      in_header: hypothesis.html
+fig_caption: yes
 bibliography: library.bib
@@
 
-__For the 2017 IFLA conference, theme: The Open Revolution__
+__For the 2017 IFLA conference, theme: Being open about open__
 
@@
 
-Scholarly communication is in dire need of disruption.
-People often tell me we need incremental change from within to produce a sustainable, modern scholarly communication system, but I oppose this idea.
-After decades of identifying various problems we have seen little sustainable change that truly and deeply reimagines how we communicate.
-Yes, PLOS changed the game.
-Yes, RIO journal is one of the most innovative ideas in scholarly communication of the last two years.
-But we are still trying to turn a horse (i.e., paper based communication) into a motor vehicle (i.e., the new thing).
+Scholarly communication is in (dire) need of disruption.
+Subscription costs to scholarly journals are untenable already [e.g., @harvard-serials-crisis] and will become even more untenable in the long run [i.e., serials crisis; @serials-crisis].
+The current scholarly incentive system detracts validity instead of adding it [e.g., @10.1371/journal.pmed.0020124;@10.1371/journal.pmed.0050201] by promoting behavior that is antithetical to the norms of science [@merton1942;@10.2307/2094423;@10.1525/jer.2007.2.4.3].
+Results cannot systematically be verified because data are rarely shared [e.g., @10.1038/nature.2017.21549;@10.1525/collabra.13;@10.1037/0003-066x.61.7.726] or preserved [@10.1016/j.cub.2013.11.014].
+Widely publicized results regularly fail reproducibility tests [e.g., @10.1126/science.aac4716;@10.1038/483531a;@10.1038/541269a, but see also @10.1016/j.jesp.2015.10.001], in part because published results are highly pre-selected for unscientific reasons [e.g., @10.1016/j.tree.2013.05.007;@10.1038/526182a;@10.1126/science.1255484;@10.1016/0140-6736(91)90201-Y
+].
+These problems are not new but they are also not absolute --- scholars want to behave in accordance with the scientific norms [@10.1525/jer.2007.2.4.3] but the system seems to encourage us to do the opposite.
 
-In this perspective piece, I try to radically reimagine the scholarly communications infrastructure based on the tools that are available to us now.
-I want these ideas to be criticized; only with diversity of ideas can we truly make progress.
-Currently the decisions are driven by a few oligopolies [@10.1371/journal.pone.0127502] or the tenured professors, the policymakers, and other people who are actually farthest removed from doing actual research.
-Many of these people stem from a different generation and have become successful in the broken system we are now stuck with.
-As a result, they might underestimate some of the problems we now face.
-This inherently decreases the diversity of ideas and stalls the development of how scholars communicate, with dire consequences (as we noticed in recent years, with unfounded and unconstructive criticisms of replication initiatives).
+In this perspective piece, I try to radically reimagine the scholarly communications infrastructure based on the digital tools that are available to us now, instead of the analog era and its legacies.
+I thoroughly believe that the Mertonian norms in science [@merton1942] align perfectly with transparency in science [see also @10.14293/s2199-1006.1.sor-socsci.arysbi.v1] and that a scholarly communications system can be built on this framework in a sustainable manner.
+I invite everyone to criticize these ideas, for only with diversity of ideas can we truly make progress.
+Currently the decisions are driven by a few oligopolies [@10.1371/journal.pone.0127502] or the highly pre-selected sample of tenured professors and policymakers, who mostly stem from the generation that has created this broken system.
+As a result, they are likely to underestimate the problems we now face and cannot be expected to intervene in the ways that are necessary.
 
 In the last decades, interpersonal communication has changed tremendously.
-These past decades have brought with it much diversity in communication, such that it can become confusing (see Figure 1).
-At the same time, the diversity is liberating by allowing innovative ways to communicate (e.g., social media) or making available previously difficult to use safety measures (e.g., encryption).
-Diversity has thus spawned new ways of communicating or have made previously esoteric communication tools available to the masses.
+Not only has it brought much diversity in communication, so much diversity has come into existence that it has become confusing (see Figure 1).
+At the same time, that same diversity has liberated us by allowing innovative ways to communicate (e.g., social media) or making available previously difficult to use safety measures easily accessible to the masses (e.g., encryption).
+Diversity in communication-channels has thus spawned new ways of communicating.
 
@@
 
-However, the way scholars communicate their findings has barely changed; we are still stuck in a paper paradigm, albeit writing on digital pages.
-In recent years, options have started to arise with tools such as [Rmarkdown]() and [Jupyter notebooks]() to break free from this paper paradigm.
-However, in the end, the scholarly communications system as we now know it still requires submitting to a journal that fits all of it back into a page-by-page, static, and paper format.
-This not only prevents new ways of communicating findings, it also prevents encouragement of discussion amongst readers about the findings, but to passively consume.
-Not to even start of the antiquated and wholly redundant process of waiting for the article to be assigned an issue number.
+However, the way scholars communicate their findings has barely changed; we are still stuck in a paper paradigm, albeit writing digital pages.
+Such digital pages radically restrict the way we imagine our communication of results.
+In recent years, options to break free from this paper paradigm have arisen with tools such as [Rmarkdown](http://rmarkdown.rstudio.com/) or [Jupyter notebooks](https://jupyter.org/).
+These options allow the user to include code to dynamically generate figures, tables, or values in the text.
+By extension, how data are cleaned and analyzed can also be included as part of a manuscript for co-authors to verify or readers to discuss [see also @elife-rmd].
+Moreover, interactive elements can be embedded, such as interactive figures, online applications, or videos.
+The current scholarly communications system still requires submitting to a journal that fits all of that back into a page-by-page, static, and paper format without all of these options.
+This bottleneck not only prevents new ways of communicating findings, it also prevents further encouragement of discussion amongst readers about the findings.
+As a result, the current system encourages scholars to passively consume findings instead of critically evaluating them.
 
-The scholarly process is much more complicated than can be captured in a static and retrospective publication.
+The scholarly process is much more complicated than can be captured in a single static and retrospective publication.
+The previously mentioned interactive elements enrich a single publication, but multiple moments of communicating ideas and results maps the empirical cycle onto the communication process.
 For example, a quantitative scholar needs to develop methods to investigate a research question.
 Feedback from peers would be valuable at that point.
-This has slowly been picked up with so called [Registered Reports]().
-However, this is more the exception than the norm.
+With so called Registered Reports [@10.1016/j.cortex.2012.12.016] a two-staged submission process is introduced to move away from a single static publication by allowing selected peers to discuss the contents at an interim stage.
+Additionally, this ensures chronology of predictions such as pre-registrations aim to do [@10.1038/s41562-016-0021], but in a more natural way.
+Currently, pre-registration and data sharing are for the proactive, hence, remain the exception rather than the norm [@10.1371/journal.pbio.1002456], while these are already implicitly part of how scholarly research is conducted.
 Additionally, researchers create materials, protocols, and other outputs during the research process that now are not (aptly) sharable in a static publication.
+If the scholarly research process has more moments of communication, these implicit parts are finally able to show themselves
 
-The scholarly process has also grown even _more_ complicated due to increased variation in the research tools available.
-Point and click software has increased the number of statistical analyses a researcher can do within minutes; the development of technology has created new types of data output that greatly exceed human comprehension and require algorithms before being able to make sense of the data (e.g., fMRI data).
-How a researcher applies all these tools is essential, but the current scholarly communication system does not require the details.
+Additionally, the scholarly process has grown more complicated with the rise of research tools available and the speed with which alternative results can be inspected.
+Point-and-click software has increased the number of statistical analyses a researcher can do within minutes.
+This increased researcher degrees of freedom to present results that are more in line with previous results, leading to confirmation bias or motivated reasoning [@10.1037/0033-2909.108.3.480], hence, invalid results [@10.1177/0956797611417632].
+Moreover, the development of technology has created new types of data output that greatly exceed human comprehension and require algorithms before being able to make sense of the data (e.g., fMRI data).
+How a researcher applies all these tools is essential, but the current scholarly communication system does not require communication of the details needed to understand what steps were taken.
 Lack of direct reproducibility of results [e.g., @10.1101/056473] therefore is no surprise when we do not know how data are processed and analyzed.
+As such, the scholarly process has become more of a black box than it was before because communication has not adapted proportionate to the increasing complexity of the message that needs to be communicated.
+Increasing complexity and lack of transparency makes it increasingly difficult to reproduce steps taken, simply due to an increasing garden of forking paths [@10.1511/2014.111.460].
 
-As such, the scholarly process has become _more_ of a black box than it was before.
-Increasing complexity and lack of transparency makes it very difficult to reproduce steps.
-It is similar to being dropped in the middle of a desert with a blindfold on, and then being asked to find your way back home.
-Moreover, much of the software is proprietary and it is simply impossible to verify results or understand how they originated [@10.1038/nature10836].
-Trying to reverse engineer these pieces of software is actually often illegal based on the unilateral agreement that you have to accept before using the software.
+<!-- Moreover, much of the software is proprietary and it is simply impossible to verify results or understand how they originated [@10.1038/nature10836].
+Trying to reverse engineer these pieces of software is actually often illegal based on the terms of service agreement that you have to accept before using the software.
 As such, it is practically impossible to understand how software development affects how results are computed, while testing for this is easy with open-source software.
-<!-- Dit moet nog herschreven worden -->
-
-Due to these problems, breaking up the reign of the paper paradigm is a necessity to better represent and understand how scholarly results come into existence.
-Of course paper is a valid medium and should not be neglected, but this is not the case when scholarly communications becomes truly digitally born.
-PDF versions of outputs can still be produced.
-
-Additionally, the way we organize scholarly communication cannot break free from the legacies if we do not radically reenvision it.
-due to the commodification of the resource (i.e., knowledge), a new system where ways to consume information is commodified are actively stifled.
-
-Commodification of consumption greatly encourages society to innovate how people can discover and reuse information.
-<!-- For Wikipedia alone, one of the greatest resources of uncommodified knowledge, there are tons of applications that allow you to consume that information in various ways on various devices.
 -->
 
-By focusing on the commodification of consumption instead of commodification of the resource, the problem of access to knowledge can be resolved in a sustainable manner.
-<!-- Politicians and scholars have been discussing Open Access for at least 15 years.
--->
+Due to the fact that these complexities currently cannot and therefore are not reported, breaking the reign of the paper paradigm is a necessity to better represent and understand how scholarly results come into existence.
+Static paper publicatoin remains a valid communication medium and it is not made obsolete when scholarly communications become truly digitally born and continuous.
+PDF versions of outputs can still be produced if a paper format is suitable or desirable.
+However, it is easy to imagine videos being embedded for protocols to be easily clarified (having "Video X" alongside "Figure X" and "Table X", for example) by origin, to then be replaced by stills when a paper version is produced.
+Nonetheless, without shifting the main focus from paper-based to digital-based scholarly communication, we cannot break free from the legacies of the paper paradigm.
+Paper-based communication includes the legacy of printing research outputs by default, which by extension includes the idea of subscriptions, purchasing content, and restricting the flow of information.
+Copyright is the legacy of the paper-based age that seems to limit the digital potential of the outputs.
+As such, the legacy of a paper-based paradigm is that of knowledge commodification.
+
+For scholarly communications, commodifying knowledge as it is currently done can be seen as a human rights violation.
+The Universal Declaration of Human Rights (UDHR), Article 27 section 1, reads as follows
+> Everyone has the right freely to participate in the cultural life of the community, to enjoy the arts and to share in scientific advancement and its benefits.
+When knowledge is commodified, it results in decreased participation, enjoyment, and sharing of scholarly progress.
+It could therefore be argued that this is a violation of human rights [@udhr].
+Moreover, this is the main legal motivation proposed by Alexandra Elbakyan as to why [Sci-Hub](http://sci-hub.bz) is not illegal from her point of view [@alexandraelbakyan].
+Legal cases on this could result in clarification of whether unauthorized sharing, which is part of scholarly culture [@10.1038/545145a], is allowed from a human rights perspective, and whether that would supercede copyright legislation.
+
+Moreover, commodifying knowledge is not the only way to make scholarly communication sustainable.
+An alternative is the commodification of how information is consumed, that is, the services that serve the users the content.
+For example, if all scholarly information is freely available, (not-)for-profit organizations can compete for users who want to sift through that information efficiently.
+If an algorithm is developed that greatly increases efficiency of finding relevant articles, revenue could be built with usage fees to that service instead of the content.
+Such a market could create competition in information consumption, which is timely given the information explosion humanity is going through at the moment.
+A free resource would encourage innovation and competition as to how people can discover and consume that information.
+
+Consequently, by focusing on the commodification of consumption instead of commodification of the resource, the problem of access to knowledge can be resolved in a sustainable manner.
+When the consumption is commodified, all service providers benefit from creating the largest pool of resources that is freely available to build their service on.
+By extension, it shifts the optimal outcome from not sharing (i.e., defecting) to sharing (i.e., cooperating), because it increases the value of their service.
+This would create a market that stimulates access instead of limiting it, while retaining the potential for revenue.
 
 Additionally, commodification of consumption removes several perverse incentives from the scholarly system that now produces unreliable knowledge.
-<!-- For example, innovativeness and surprisingness are criteria upon which articles are currently selected, where those exact some properties increase the probability that the finding is false [@10.1371/journal.pmed.0020124].
--->
+For example, innovativeness and surprisingness are criteria upon which articles are currently selected, where those exact some properties increase the probability that the finding is false [@10.1371/journal.pmed.0020124].
+When the knowledge is commodified, this makes sense from a business perspective (comparable to news outlets): readers will buy access more to shocking and astounding results.
+However, when consumption is commodified, this whole incentive is removed and the most valuable resource is the most populated and diverse one.
+This would allow much more information to be extracted, hence, richer services to be built on top of that resource.
 
-The main tenet underlying the themes of Open Access, Open Data, Open Science, and replication initiatives in scholarly communication is sustainability.
-<!-- Sustainability of consuming knowledge (Open Access), sustainability of producing reliable knowledge ("reproducibility crisis"), sustainability of The serial crisis indicates that access to knowledge is currently unsustainable (e.g., Harvard cannot even access everything with its massive budget).
--->
-
-In this perspective piece, I propose a redesign of the scholarly communication system that is flexible, decentralized, and distributed, which immediately allows for more sustainability and diversity in the way we consume, produce, and access knowledge.
+The main tenet underlying the themes of Open Access, Open Data, Open Science, and replication initiatives in scholarly communication is sustainability through transparency of the scholarly process in all facets.
+As I tried to explain in the previous paragraphs, this sustainability can be achieved by recalibrating the fundamental business model of the scholarly communications system.
+In this perspective piece, I propose a redesign of the scholarly communication system that is flexible, decentralized, distributed, and freely reusable from the start.
+Such a system would allow for more sustainability and diversity in the way we consume, produce, and access knowledge, which would ultimately benefit everyone who reaps the benefits of scholarly results.
 
@@
 
-The scholarly process is ridden with single or few points of failures and sharing makes this a many points for failure system.
+The scholarly process is ridden with single or few points of failures at the moment.
 <!-- For exactly this reason it is so important that scholars become more transparent about their data, their hypotheses, their results, and the chronology of a.
 -->
+
+These points of failures occur at each stage of the scholarly process.
 
@@
 
-Hence, a sustainable scholarly communications system needs to be both decentralized and distributed.
+In order to remove these failure points, a sustainable scholarly communications system would have to be both decentralized and widely distributed.
 
@@
 
-Alongside decentralization and distribution, integrity of the scholarly record can also become better.
+Alongside decentralization and distribution, integrity of the scholarly record can also improve.
 The chronology of the scholarly process can be ensured by splitting up the paper article as we know it into nodes of information.
 
-These nodes can easily be logged on a blockchain based ledger of which everyone can have a copy (e.g., [Hyperledger](https://github.com/hyperledger/fabric)-based).
-<!-- An example is available below -->
+These nodes can easily be logged on a blockchain based ledger (e.g., [Hyperledger](https://github.com/hyperledger/fabric)-based) of which everyone can have a copy.
+Although explaining the entire operating principle of the blockchain is beyond the scope of this piece, the essence is very simple.
+Each entry on the blockchain needs to satisfy a predefined mathematical rule [@], which is dependent on the contents of the previous entry in the ledger, the contents of the new entry, and a constant to offset the current entry.
+Solving this offset for the new entry to satisfy a specific mathematical rule is called 'mining'.
+However, because the proof of record is dependent on the contents of the previous entry (i.e., block), the chain is broken when the contents of that previous entry are changed.
+Given that each user has a copy of the entire blockchain, those who have a different chain from the majority of other users are simply rejected.
+As such, in order to change the ledger, each change would have to be re-verified by the entire network to become verified (for an introduction to the blockchain, see https://www.youtube.com/watch?v=_160oMzblY8).
 
@@
 	"parent": "dat://92b9b1b95e18c1dd1546c25ac49609da6fb0e3f2abc94ecf17609c565888c16b",
-	"self": "dat://c2c291f93fa635c7fc7bc6e8ba504937b1821c5be45ab7c859f6e6cd19a291b7",
+	"self": "/ipfs/QmVLDAhCY3X9P2uRudKAryuQFPM5zqA3Yij1dY8FpGbL7",
 	"time-deposit": "2017-04-12T12:07:06+02:00",
@@
 
-Each entry in the ledger needs to be self-containing and refer to all relevant files.
+Each entry in the ledger needs to be self-containing and refer to all relevant files. The ledger can be platform agnostic.
 
@@
 
-The ledger can also allow for time-locked sharing but with a clear chronology of creation.
+The ledger can also allow for time-locked sharing but with a clear chronology of creation. Time-locked sharing would require a third-party that can be trusted
 
@@
 
-Combinations of ORCIDs can be hashed into unique author combinations (and ordering!) to provide identification over time, but remove gender and status bias (amongst other things).
+Combinations of ORCIDs can be hashed into unique author combinations to provide identification over time, but remove gender and status bias during initial assessment (amongst other things).
 

### 2017-05-23 manuscript.Rmd
Full draft!

@@
       in_header: hypothesis.html
+  word_document: default
 fig_caption: yes
@@
 
+__Made available under a [CC 0 public domain dedication](https://creativecommons.org/publicdomain/zero/1.0/legalcode)__
+
 ## Abstract
+
+Scholarly communication is in need of disruption.
+Commodifying knowledge as is currently done with journals, is not sustainable any longer.
+An alternative is the commodification of how information is consumed.
+By focusing on the commodification of consumption instead of commodification of the resource, the problem of access to knowledge can be resolved in a sustainable manner.
+Additionally, commodification of consumption removes several perverse incentives from the scholarly system that now produces unreliable knowledge.
+The main tenet underlying the themes of Open Access, Open Data, Open Science, and replication initiatives in scholarly communication is sustainability through transparency of the scholarly process in all facets.
+The sustainability of any networked system is threatened by single points of failure (i.e., the entire system can be manipulated from one node in the network).
+The scholarly process is ridden with such single points of failures at all stages.
+Distributing the scholarly communications system would remove the problems of single points of failure.
+Distributing and decentralizing the scholarly communications system is achievable with newly developed peer-to-peer (p2p) Internet protocols.
+Alongside decentralization and distribution of the content, integrity of the scholarly record can also be reformed to transform sections of a paper into different, reusable nodes of knowledge.
+These nodes can be logged on a blockchain based ledger of which everyone can have a copy.
+In order to deposit nodes onto the ledger, the depositor needs to agree that the contents are licensed CC 0, in order to maximize legal certainty regarding reuse of the contents.
+This is key to create a sustainable eco-system where scholars and companies can cooperate instead of compete, as we currently do.
 
@@
 Scholarly communication is in (dire) need of disruption.
-Subscription costs to scholarly journals are untenable already [e.g., @harvard-serials-crisis] and will become even more untenable in the long run [i.e., serials crisis; @serials-crisis].
+Subscription costs to scholarly journals are untenable already [e.g., @harvard-serials-crisis] and will become even more untenable in the long run with above inflation price-increases [i.e., serials crisis; @serials-crisis].
 The current scholarly incentive system detracts validity instead of adding it [e.g., @10.1371/journal.pmed.0020124;@10.1371/journal.pmed.0050201] by promoting behavior that is antithetical to the norms of science [@merton1942;@10.2307/2094423;@10.1525/jer.2007.2.4.3].
-Results cannot systematically be verified because data are rarely shared [e.g., @10.1038/nature.2017.21549;@10.1525/collabra.13;@10.1037/0003-066x.61.7.726] or preserved [@10.1016/j.cub.2013.11.014].
-Widely publicized results regularly fail reproducibility tests [e.g., @10.1126/science.aac4716;@10.1038/483531a;@10.1038/541269a, but see also @10.1016/j.jesp.2015.10.001], in part because published results are highly pre-selected for unscientific reasons [e.g., @10.1016/j.tree.2013.05.007;@10.1038/526182a;@10.1126/science.1255484;@10.1016/0140-6736(91)90201-Y
-].
+For example, results particular to one researcher instead of universal among researchers, because results cannot systematically be verified considering data are rarely shared [@10.1038/nature.2017.21549;@10.1525/collabra.13;@10.1037/0003-066x.61.7.726] or preserved [@10.1016/j.cub.2013.11.014].
+Widely publicized results regularly fail reproducibility tests [e.g., @10.1126/science.aac4716;@10.1038/483531a;@10.1038/541269a; but see also @10.1016/j.jesp.2015.10.001], in part because published results are highly pre-selected for unscientific reasons [e.g., pure novelty of results, @10.1016/j.tree.2013.05.007; easy to assimilate findings, @10.1038/526182a; statistical significance, @10.1126/science.1255484;@10.1016/0140-67369190201-Y].
 These problems are not new but they are also not absolute --- scholars want to behave in accordance with the scientific norms [@10.1525/jer.2007.2.4.3] but the system seems to encourage us to do the opposite.
@@
 I invite everyone to criticize these ideas, for only with diversity of ideas can we truly make progress.
-Currently the decisions are driven by a few oligopolies [@10.1371/journal.pone.0127502] or the highly pre-selected sample of tenured professors and policymakers, who mostly stem from the generation that has created this broken system.
-As a result, they are likely to underestimate the problems we now face and cannot be expected to intervene in the ways that are necessary.
+Currently, how scholars operate (i.e., governance) is driven by a few oligopolies [@10.1371/journal.pone.0127502] or the highly pre-selected sample of tenured professors and policymakers, who mostly stem from the generation that has created this broken system and have benefited or still benefit from the current system.
+As a result, they are likely to underestimate the need for change in scholarly communications and cannot be expected to intervene in the necessary ways.
 
@@
 
-```{r xkcd-1810, fig.cap="_Figure 1._ A comical example of the diversity in communication channels.", out.width="60%", fig.align="center", echo = FALSE}
+```{r xkcd-1810, fig.cap="_Figure 1._ A comical example of the diversity in communication channels.
+[Licensed under CC BY-NC 2.5 from xkcd](https://xkcd.com/1810/)", out.width="60%", fig.align="center", echo = FALSE}
 knitr::include_graphics('figs/xkcd_1810-cc-by-nc-2_5.png', auto_pdf = TRUE)
@@
 
-However, the way scholars communicate their findings has barely changed; we are still stuck in a paper paradigm, albeit writing digital pages.
-Such digital pages radically restrict the way we imagine our communication of results.
+The current scholarly communications system requires submitting to a journal that fits all scholarly output into a static, page-by-page, and paper format without the innovative solutions that are now available.
+This bottleneck not only prevents new ways of communicating findings, it also prevents further encouragement of discussion amongst readers about the findings.
+As a result, the current system encourages scholars to passively consume findings instead of critically evaluating them.
+
+Hence, we are stuck in a paper paradigm, albeit writing digital pages.
+Such digital pages radically restrict the way we imagine the communication of our results.
 In recent years, options to break free from this paper paradigm have arisen with tools such as [Rmarkdown](http://rmarkdown.rstudio.com/) or [Jupyter notebooks](https://jupyter.org/).
@@
 Moreover, interactive elements can be embedded, such as interactive figures, online applications, or videos.
-The current scholarly communications system still requires submitting to a journal that fits all of that back into a page-by-page, static, and paper format without all of these options.
-This bottleneck not only prevents new ways of communicating findings, it also prevents further encouragement of discussion amongst readers about the findings.
-As a result, the current system encourages scholars to passively consume findings instead of critically evaluating them.
 
@@
 Feedback from peers would be valuable at that point.
-With so called Registered Reports [@10.1016/j.cortex.2012.12.016] a two-staged submission process is introduced to move away from a single static publication by allowing selected peers to discuss the contents at an interim stage.
+With so called Registered Reports [@10.1016/j.cortex.2012.12.016] a two-staged submission process is introduced to move away from a single static publication.
+It allows selected peers (i.e., peer reviewers) to discuss the contents at an interim stage.
 Additionally, this ensures chronology of predictions such as pre-registrations aim to do [@10.1038/s41562-016-0021], but in a more natural way.
-Currently, pre-registration and data sharing are for the proactive, hence, remain the exception rather than the norm [@10.1371/journal.pbio.1002456], while these are already implicitly part of how scholarly research is conducted.
-Additionally, researchers create materials, protocols, and other outputs during the research process that now are not (aptly) sharable in a static publication.
-If the scholarly research process has more moments of communication, these implicit parts are finally able to show themselves
+Nonetheless, registered reports, pre-registration, and data sharing are for the proactive, hence, remain the exception rather than the norm [@10.1371/journal.pbio.1002456], while these are already implicitly part of how scholarly research is conducted.
+Additionally, researchers create materials, protocols, and other outputs during the research process, share them with colleagues/advisors, but this process is not (aptly) sharable in a static publication.
+If the scholarly research process has more moments of communication, these implicit parts are finally able to show themselves.
 
@@
 Due to the fact that these complexities currently cannot and therefore are not reported, breaking the reign of the paper paradigm is a necessity to better represent and understand how scholarly results come into existence.
-Static paper publicatoin remains a valid communication medium and it is not made obsolete when scholarly communications become truly digitally born and continuous.
+Static paper publication remains a valid communication medium and it is not made obsolete when scholarly communications become truly digitally born and continuous.
 PDF versions of outputs can still be produced if a paper format is suitable or desirable.
@@
 Paper-based communication includes the legacy of printing research outputs by default, which by extension includes the idea of subscriptions, purchasing content, and restricting the flow of information.
-Copyright is the legacy of the paper-based age that seems to limit the digital potential of the outputs.
-As such, the legacy of a paper-based paradigm is that of knowledge commodification.
+Copyright is the legacy of the paper-based age that seems to limit the digital potential of scholarly outputs.
+As such, the legacy of a paper-based paradigm with respect to scholarly communication is that of knowledge commodification.
 
-For scholarly communications, commodifying knowledge as it is currently done can be seen as a human rights violation.
-The Universal Declaration of Human Rights (UDHR), Article 27 section 1, reads as follows
+For scholarly communication, commodifying knowledge as it is currently done can be seen as a human rights violation.
+[The Universal Declaration of Human Rights (UDHR), Article 27 section 1](http://wayback.archive.org/web/20170510094028/http://www.un.org/en/universal-declaration-human-rights/index.html), reads as follows
+
 > Everyone has the right freely to participate in the cultural life of the community, to enjoy the arts and to share in scientific advancement and its benefits.
+
 When knowledge is commodified, it results in decreased participation, enjoyment, and sharing of scholarly progress.
@@
 Moreover, this is the main legal motivation proposed by Alexandra Elbakyan as to why [Sci-Hub](http://sci-hub.bz) is not illegal from her point of view [@alexandraelbakyan].
-Legal cases on this could result in clarification of whether unauthorized sharing, which is part of scholarly culture [@10.1038/545145a], is allowed from a human rights perspective, and whether that would supercede copyright legislation.
+Discussion on the legal correctness of this argument would be worthwhile to further learn about this.
+On the other hand, legal cases on this could result in clarification of whether unauthorized sharing, which is part of scholarly culture [@10.1038/545145a], is allowed from a human rights perspective and whether such sharing would supercede copyright legislation.
 
-Moreover, commodifying knowledge is not the only way to make scholarly communication sustainable.
-An alternative is the commodification of how information is consumed, that is, the services that serve the users the content.
+Moreover, commodifying knowledge is not the only way to make scholarly communication sustainable; an alternative is the commodification of how information is consumed.
+That is, commodifying the services that serve the users the content, instead of the content itself.
 For example, if all scholarly information is freely available, (not-)for-profit organizations can compete for users who want to sift through that information efficiently.
@@
 Such a market could create competition in information consumption, which is timely given the information explosion humanity is going through at the moment.
-A free resource would encourage innovation and competition as to how people can discover and consume that information.
+The potential market capitalization of free flow of information and data is tremendous [e.g., freely reusable public sector information/data in the U.K.
+is estimated to provide a market cap of £590 million - £16 billion; p.
+96 of @dotecon].
+A freely reusable information resource would provide the infrastructure that encourages innovation and competition as to how people can discover and consume that information.
 
@@
 As I tried to explain in the previous paragraphs, this sustainability can be achieved by recalibrating the fundamental business model of the scholarly communications system.
-In this perspective piece, I propose a redesign of the scholarly communication system that is flexible, decentralized, distributed, and freely reusable from the start.
+In the next section of this perspective piece, I propose a redesign of the scholarly communication system that is flexible, decentralized, distributed, and freely accessible and reusable from the start.
 Such a system would allow for more sustainability and diversity in the way we consume, produce, and access knowledge, which would ultimately benefit everyone who reaps the benefits of scholarly results.
 
-## Sustainable scholarly process
+## Sustainable scholarly communication
 
-Single points of failures make a system vulnerable to outages, (malicious) adjustment of content, or even complete removal of content.
-<!-- As such, a centralized infrastructure is easier to disrupt than a decentralized or distributed one.
-The Internet is currently organized in such a way that it is relatively centralized, whereas the Darknet (i.e., an alternative cyberspace) is less centralized and has been found to be more resilient [@10.1103/PhysRevE.95.022313].
--->
+The sustainability of any networked system is threatened by single points of failure (i.e., the entire system can be manipulated from one node in the network).
+Single points of failure not only make a system vulnerable to downtime, it also raises the possibility of (malicious) adjustment of content, or even complete removal of content (among other things).
+For example, retracted papers sometimes wholly disappear, leaving only the retraction notice; this results in a direct and nearly irreversable adjustment of the scholarly record (from the scholar's perspective).
+Due to these single (or few) points of failure, such a centralized and concentrated infrastructure is easier to disrupt than a decentralized or distributed one [see also @10.1103/physreve.95.022313].
 
-The scholarly process is ridden with single or few points of failures at the moment.
-<!-- For exactly this reason it is so important that scholars become more transparent about their data, their hypotheses, their results, and the chronology of a.
--->
+The scholarly process is ridden with single- or few points of failures at all stages.
+Hence its sustainability is directly threatened.
+As mentioned before, communication occurs only once (i.e., at the end); this could be regarded as a single point of failure for communicating what the scholars did.
+If data are not shared or preserved, the scholar creates a single point of failure in accessing and preserving the data.
+When results are published with a closed access publisher, the publisher becomes the single point of failure for providing access to those results.
+For exactly this reason I find it important that scholars become more transparent about their data, their hypotheses, their results, and the chronology of their research process; it decreases the risk of a single point of failure actually failing.
 
-These points of failures occur at each stage of the scholarly process.
+Distributing the scholarly communications system would remove the problems of single- or few points of failure in access to findings.
+Currently, [Portico](http://www.portico.org/digital-preservation/) and [CLOCKSS](https://clockss.org/clockss/Home) are supposed to provide access in case the original content is subject to a ["trigger event"](https://clockss.org/clockss/FAQ#How_does_the_CLOCKSS_board_define_a_trigger_event.3F) (e.g., publisher goes out of business).
+However, this system is insufficient for wide distribution because content is only selectively distributed to libraries.
+Moreover, by its nature "trigger events" are reactive and need governance (indicated by the Centralized LOCKSS system, i.e., CLOCKSS).
+That such governance is problematic, is subscribed by the case that CLOCKSS has previously preserved content under a more restrictive license than was originally used ([CC BY-NC-ND for the preserved journal "Tijdschrift voor Tijdschriftstudies"](http://wayback.archive.org/web/20170518110938/https://clockss.org/clockss/Tijdschrift_voor_Tijdschriftstudies), whereas [the original was CC BY](http://triggered.stanford.clockss.org/ServeContent?url=https%3A%2F%2Fwww.tijdschriftstudies.nl%2Farticles%2F10.18352%2Fts.340%2Fprint%2F)).
+In order to make a truly distributed system sustainable, persistent nodes in the network need to be identified to store all this information (e.g., libraries), but institutions and individuals should be able to freely participate in the distributed network by freely creating temporary or persistent copies for wide and unlimited distribution of content.
 
-<!-- The scholarly communications system is currently centralized at publisher websites, is insecure, and is not resilient.
-HTTP(S) because redistribution is frequently by copyright Unrestrictive licensing, such as the Creative Commons BY license, allow for free redistribution and increases persistence.
-For example, anyone can download the entire corpus of Hindawi right now and if Hindawi would then become inaccessible, the articles would still be available regardless.
-Copyright creates dependence whereas the lack thereof creates independence.
--->
+Distributing and decentralizing the scholarly communications system is achievable with peer-to-peer (p2p) Internet protocols such as [`dat`](https://datproject.org/) and [`ipfs`](https://ipfs.io/).
+Simply put, such p2p networks securely send information across a network of peers but are resilient to nodes being removed or adjusted because they operate in a mesh network.
+For example, if 20 peers have file X, removing one peer does not affect the availability of the file X.
+Only if all 20 are removed from the network, file X will become unavailable.
+Vice versa, if more peers on the network have file X, it is less likely that file X will become unavailable.
+As such, this would include unlimited redistribution in the scholarly communication system by default, instead of limited redistribution due to copyright as it is now.
 
-In order to remove these failure points, a sustainable scholarly communications system would have to be both decentralized and widely distributed.
+Alongside decentralization and distribution of the content as described in the previous paragraphs, integrity of the scholarly record can also be reformed to transform sections of a paper into different, reusable nodes of knowledge.
+The chronology of the scholarly process can be ensured by splitting up the paper article as we know it into nodes of information.
+The theory section could be a node, from which a hypotheses node follows, from which a methods node follows, and so on.
+This would provide a more true depiction of the scholarly process and facilitate direct replications (one needs to only take the materials node and create a new data node), reanalyses (one needs to only take the data node and create a new results node), etc.
+Additionally, scholars would be able to start citing more granularly by referring to specific nodes instead of entire papers and would not have to write the same theory section (for example) multiple times for different papers.
 
-Decentralization is achievable with new peer-to-peer (p2p) Internet protocols such as `dat` and `ipfs`.
-<!-- Such p2p networks send information across the network, but are is resilient to nodes being removed because they operate in mesh network.
-If 20 peers have the files, removing one from the network does not affect the availability of the files-->
+These nodes can be logged on a blockchain based ledger (e.g., [Hyperledger](https://github.com/hyperledger/fabric)-based) of which everyone can have a copy.
+Although explaining the entire operating principle of the blockchain is beyond the scope of this piece, the essence is relatively simple.
+Each entry on the blockchain needs to satisfy a predefined mathematical rule, which is dependent on the contents of the previous entry in the ledger, the contents of the new entry, and a constant to offset the current entry [@bitcoin].
+Creating a situation that fullfills that mathematical rule requires solving for the offset in the new entry to the blockchain, which is called 'mining'.
+Because the proof of record (i.e., satisfying the predefined mathematical rule) is dependent on the combination between the contents of the previous- and current entries (i.e., blocks) and the offset, the chain is broken when the contents of any entry in that chain are (unwillingly) changed.
+Given that each user has a copy of the entire blockchain, those who have a different chain from the majority of other users are simply rejected.
+As such, in order to change the ledger, each change would have to be re-verified by the majority of the network (i.e., >50%) to become verified (for a good introduction to the blockchain, see [https://www.youtube.com/watch?v=_160oMzblY8](https://www.youtube.com/watch?v=_160oMzblY8)).
 
-Such p2p networks also allow for ready redistribution when copyright is leveraged to work for distribution.
+Figure 2 shows an example of how the nodes might be incorporated into such a blockchain based ledger.
+The idea behind this format is that it needs to contain the most crucial information, but also be flexible to accompany different demands in different scholarly fields and development in scholarly research over time.
+The "parent" entry indicates what the related nodes are, which chronologically predate the new node.
+This is not meant to replace citations, but only to indicate direct links between nodes if applicable.
+For example, there might be a "data" node that spawns competing research teams to analyze the data in different ways, creating two "results" nodes that come from the same "data" node.
+Subsequently, each node can itself become a parent at a later point, which is why the "self" entry is included.
+The "parent" and "self" can be any p2p address from `dat`, `ipfs` or future p2p platform (i.e., this approach is platform agnostic).
+This efficiently combines the reference to the previous nodes and its contents, and the contents of the new node.
 
-Alongside decentralization and distribution, integrity of the scholarly record can also improve.
-The chronology of the scholarly process can be ensured by splitting up the paper article as we know it into nodes of information.
-
-These nodes can easily be logged on a blockchain based ledger (e.g., [Hyperledger](https://github.com/hyperledger/fabric)-based) of which everyone can have a copy.
-Although explaining the entire operating principle of the blockchain is beyond the scope of this piece, the essence is very simple.
-Each entry on the blockchain needs to satisfy a predefined mathematical rule [@], which is dependent on the contents of the previous entry in the ledger, the contents of the new entry, and a constant to offset the current entry.
-Solving this offset for the new entry to satisfy a specific mathematical rule is called 'mining'.
-However, because the proof of record is dependent on the contents of the previous entry (i.e., block), the chain is broken when the contents of that previous entry are changed.
-Given that each user has a copy of the entire blockchain, those who have a different chain from the majority of other users are simply rejected.
-As such, in order to change the ledger, each change would have to be re-verified by the entire network to become verified (for an introduction to the blockchain, see https://www.youtube.com/watch?v=_160oMzblY8).
-
-```json
-{
-	"parent": null,
-	"self": "dat://92b9b1b95e18c1dd1546c25ac49609da6fb0e3f2abc94ecf17609c565888c16b",
-	"time-deposit": "2017-04-10T18:38:03+02:00",
-	"time-available": "2017-04-10T18:38:03+02:00",
-	"authors": "b601f276870ebf4920b3c78c4af51d3ed3bf4db7ab2b9c4faa74e0b457aa13ed",
-	"type": "hypotheses"
-}
-
-{
-	"parent": "dat://92b9b1b95e18c1dd1546c25ac49609da6fb0e3f2abc94ecf17609c565888c16b",
-	"self": "/ipfs/QmVLDAhCY3X9P2uRudKAryuQFPM5zqA3Yij1dY8FpGbL7",
-	"time-deposit": "2017-04-12T12:07:06+02:00",
-	"time-available": "2017-08-12T12:07:06+02:00",
-	"authors": "b601f276870ebf4920b3c78c4af51d3ed3bf4db7ab2b9c4faa74e0b457aa13ed",
-	"type": "data"
-}
+```{r fig_json, fig.cap="_Figure 2._ An example of how two nodes might be described and stored in the blockchain and a depiction of how these two entries relate.", echo=FALSE, out.width="100%"}
+knitr::include_graphics('figs/fig-json.png')
 ```
 
-Each entry in the ledger needs to be self-containing and refer to all relevant files. The ledger can be platform agnostic.
+The nodes also include "time-deposit" and "time-available" entries; these indicate the time-related aspects of the node.
+To ensure chronology, the time the node is deposited into the ledger is included, ensuring that if used as a parent, it precedes the child node.
+Additionally, it makes sense that not all information can be made available at the same time as the deposit takes place, which is why "time-available" is included to make clear when that information will become available.
+There are various ways in which this time-locked system could be implemented.
+Either the author can bear the responsibility to deposit the materials in a timely manner, or a trusted third-party can provide a service to automatically deposit these materials.
+Alternatively, files could be encrypted and deposited at the same time as the entry to the ledger is made, where the private key needed to decrypt the files could be provided at the "time-available".
+A system that incorporates "time-available" moves the discussion of whether materials should be shared at all to when they can be shared (i.e., not much can be justified to stay unavailable into perpetuity).
 
-A parent node can never occur after a child node,
+In order to provide identification and accountability of nodes, author ORCIDs can be combined into unique author combinations, which also diminishes gender and status bias during initial assessment (amongst other things).
+These combinations can be created based on so-called hashes (e.g., SHA256SUM with salting).
+For example, my ORCID is [`0000-0003-1050-6809`](http://orcid.org/0000-0003-1050-6809) and that of one of my supervisors is [`0000-0003-2415-2933`](http://orcid.org/0000-0003-2415-2933).
+If I am the sole author, my "authors" hash would be ``r digest::digest('0000-0003-1050-6809', 'sha256')``, but if I am first author and my supervisor is second author, the hash is different (i.e., ``r digest::digest('0000-0003-1050-6809;0000-0003-2415-2933', 'sha256')``).
+If we do not disclose these hashes or that we authored these nodes in some composition, it is practically impossible to reverse the hash into the identifiable ORCIDs.
+These hashes are not meant as a foolproof way to safeguard author identity and remove biases in an absolute manner; these hashes are primarily meant as a way to reduce the size of each entry in the ledger (i.e., the hash is of the same length regardless of how many authors are included) and to obfuscate the author's identity.
+Such obfuscation is helpful to reduce bias in peer assessments.
 
-The ledger can also allow for time-locked sharing but with a clear chronology of creation. Time-locked sharing would require a third-party that can be trusted
+What types of nodes are contained in the ledger is up for discussion.
+For example, it is relatively clear that there can be "theory", "hypothesis", "methods", "data", "analysis", "results", and "discussion" nodes to cover the largest part of empirical research.
+However, there are various other types of nodes that will not fit into this rather restrictive taxonomy (e.g., "opinion" or those related to qualitative research).
+Moreover, over time, new types of nodes might be the result of natural development in scholarly research (e.g., due to additions or changes to the scientific method).
+As such, the system needs to be dynamic about the types of nodes it can include, but there should also not be total chaos in the types of nodes.
+This node taxonomy will require empirical study within the scholarly community to provide an evidence-base for an initial version.
 
-Nodes cannot be made available
+The contents of each node can vary, depending on the type of node.
+For example, one node can refer to just one file directly (e.g., a `csv` dataset), whereas others can refer to entire folders, containing sets of files.
+For example, a data node can contain multiple data files if it refers to a folder, or a theory section can contain the text in various (open) formats, such as Markdown and XML.
+Moreover, nodes on p2p networks such as `dat` by default support large files (e.g., 50GB); such large file storage is becoming increasingly necessary with the large amount of data available.
 
-Combinations of ORCIDs can be hashed into unique author combinations to provide identification over time, but remove gender and status bias during initial assessment (amongst other things).
+In order to deposit nodes onto the ledger, the depositor needs to agree that the contents are licensed CC 0, in order to maximize legal certainty regarding reuse of the contents.
+The current copyright system has created the problem of inaccessibility and CC 0 empowers access by default.
+Moreover, restrictive licensing also comes with duties to enforce the license, if it is to mean anything to begin with.
+Enforcing a restrictive license requires scholars to legally pursue those who infringe their copyrights, detracting from their research time.
+As such, scholars are not the ones who benefit from restrictive licensing, and neither does the scholarly community at large.
 
-What types of nodes are contained in the ledger, is something that requires discussion and can change over time. Nonetheless, the fact remains that if someone presents results without
+# Conclusion
 
-For a sustainable distribution system, persistent nodes need to be identified: libraries.
-
-Nonetheless, other stakeholders benefiting from maintaining the distribution of the decentralized network are (not-)for-profit companies that build services on top of the resources.
-
-## Locking the system open
-
-When Brewster Kahle spoke at OpenCon2016, he mentioned that we need to lock the system open.
-At first, it would seem only the open advocates win in that case, but I think the for-profit companies can benefit from this as well by opening up a whole new market where more return can be made and in a more sustainable manner.
+At OpenCon2016, Brewster Kahle mentioned that the scholarly system should be "locked open".
+After this talk I spent much time thinking about how this could be done.
+Using a decentralized and distributed system, as I tried to conceptualize throughout this article, is my initial attempt at realizing a "locked open" system that benefits not just the scholarly community, but also those that aim to generate value from these outputs.
+By shifting from a knowledge commodification to a commodification of how that knowledge is consumed, free access and reuse becomes beneficial to all parties.
+This is key to create a sustainable eco-system where scholars and companies can cooperate instead of compete, as we currently do.
 

### 2017-05-28 manuscript.Rmd
Update draft

@@
 output:
+  pdf_document: default
   html_document:

### 2017-05-28 manuscript.tex
Update draft

--- 2017-05-23 manuscript.Rmd
@@
----
-title: "Re-envisioning a future in scholarly communication"
-author: "Chris H.J. Hartgerink"
-date: '`r Sys.Date()`'
-output:
-  html_document:
-    includes:
-      in_header: hypothesis.html
-  word_document: default
-fig_caption: yes
-bibliography: library.bib
----
+\documentclass[]{tufte-book}
 
-__For the 2017 IFLA conference, theme: Being open about open__
+% ams
+\usepackage{amssymb,amsmath}
 
-__Made available under a [CC 0 public domain dedication](https://creativecommons.org/publicdomain/zero/1.0/legalcode)__
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \makeatletter
+  \@ifpackageloaded{fontspec}{}{\usepackage{fontspec}}
+  \makeatother
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \makeatletter
+  \@ifpackageloaded{soul}{
+     \renewcommand\allcapsspacing[1]{{\addfontfeature{LetterSpace=15}#1}}
+     \renewcommand\smallcapsspacing[1]{{\addfontfeature{LetterSpace=10}#1}}
+   }{}
+  \makeatother
+\fi
 
-## Abstract
+% graphix
+\usepackage{graphicx}
+\setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}
 
-Scholarly communication is in need of disruption.
-Commodifying knowledge as is currently done with journals, is not sustainable any longer.
-An alternative is the commodification of how information is consumed.
-By focusing on the commodification of consumption instead of commodification of the resource, the problem of access to knowledge can be resolved in a sustainable manner.
-Additionally, commodification of consumption removes several perverse incentives from the scholarly system that now produces unreliable knowledge.
-The main tenet underlying the themes of Open Access, Open Data, Open Science, and replication initiatives in scholarly communication is sustainability through transparency of the scholarly process in all facets.
-The sustainability of any networked system is threatened by single points of failure (i.e., the entire system can be manipulated from one node in the network).
-The scholarly process is ridden with such single points of failures at all stages.
-Distributing the scholarly communications system would remove the problems of single points of failure.
-Distributing and decentralizing the scholarly communications system is achievable with newly developed peer-to-peer (p2p) Internet protocols.
-Alongside decentralization and distribution of the content, integrity of the scholarly record can also be reformed to transform sections of a paper into different, reusable nodes of knowledge.
-These nodes can be logged on a blockchain based ledger of which everyone can have a copy.
-In order to deposit nodes onto the ledger, the depositor needs to agree that the contents are licensed CC 0, in order to maximize legal certainty regarding reuse of the contents.
-This is key to create a sustainable eco-system where scholars and companies can cooperate instead of compete, as we currently do.
+% booktabs
+\usepackage{booktabs}
 
-## Body
+% url
+\usepackage{url}
 
-Scholarly communication is in (dire) need of disruption.
-Subscription costs to scholarly journals are untenable already [e.g., @harvard-serials-crisis] and will become even more untenable in the long run with above inflation price-increases [i.e., serials crisis; @serials-crisis].
-The current scholarly incentive system detracts validity instead of adding it [e.g., @10.1371/journal.pmed.0020124;@10.1371/journal.pmed.0050201] by promoting behavior that is antithetical to the norms of science [@merton1942;@10.2307/2094423;@10.1525/jer.2007.2.4.3].
-For example, results particular to one researcher instead of universal among researchers, because results cannot systematically be verified considering data are rarely shared [@10.1038/nature.2017.21549;@10.1525/collabra.13;@10.1037/0003-066x.61.7.726] or preserved [@10.1016/j.cub.2013.11.014].
-Widely publicized results regularly fail reproducibility tests [e.g., @10.1126/science.aac4716;@10.1038/483531a;@10.1038/541269a; but see also @10.1016/j.jesp.2015.10.001], in part because published results are highly pre-selected for unscientific reasons [e.g., pure novelty of results, @10.1016/j.tree.2013.05.007; easy to assimilate findings, @10.1038/526182a; statistical significance, @10.1126/science.1255484;@10.1016/0140-67369190201-Y].
-These problems are not new but they are also not absolute --- scholars want to behave in accordance with the scientific norms [@10.1525/jer.2007.2.4.3] but the system seems to encourage us to do the opposite.
+% hyperref
+\usepackage{hyperref}
 
-In this perspective piece, I try to radically reimagine the scholarly communications infrastructure based on the digital tools that are available to us now, instead of the analog era and its legacies.
-I thoroughly believe that the Mertonian norms in science [@merton1942] align perfectly with transparency in science [see also @10.14293/s2199-1006.1.sor-socsci.arysbi.v1] and that a scholarly communications system can be built on this framework in a sustainable manner.
-I invite everyone to criticize these ideas, for only with diversity of ideas can we truly make progress.
-Currently, how scholars operate (i.e., governance) is driven by a few oligopolies [@10.1371/journal.pone.0127502] or the highly pre-selected sample of tenured professors and policymakers, who mostly stem from the generation that has created this broken system and have benefited or still benefit from the current system.
-As a result, they are likely to underestimate the need for change in scholarly communications and cannot be expected to intervene in the necessary ways.
+% units.
+\usepackage{units}
 
-In the last decades, interpersonal communication has changed tremendously.
-Not only has it brought much diversity in communication, so much diversity has come into existence that it has become confusing (see Figure 1).
-At the same time, that same diversity has liberated us by allowing innovative ways to communicate (e.g., social media) or making available previously difficult to use safety measures easily accessible to the masses (e.g., encryption).
-Diversity in communication-channels has thus spawned new ways of communicating.
+\setcounter{secnumdepth}{-1}
 
-```{r xkcd-1810, fig.cap="_Figure 1._ A comical example of the diversity in communication channels.
-[Licensed under CC BY-NC 2.5 from xkcd](https://xkcd.com/1810/)", out.width="60%", fig.align="center", echo = FALSE}
-knitr::include_graphics('figs/xkcd_1810-cc-by-nc-2_5.png', auto_pdf = TRUE)
-```
+% citations
+\usepackage{natbib}
+\bibliographystyle{plainnat}
 
-The current scholarly communications system requires submitting to a journal that fits all scholarly output into a static, page-by-page, and paper format without the innovative solutions that are now available.
-This bottleneck not only prevents new ways of communicating findings, it also prevents further encouragement of discussion amongst readers about the findings.
-As a result, the current system encourages scholars to passively consume findings instead of critically evaluating them.
+% pandoc syntax highlighting
+
+% longtable
+
+% multiplecol
+\usepackage{multicol}
+
+% strikeout
+\usepackage[normalem]{ulem}
+
+% morefloats
+\usepackage{morefloats}
+
+% tightlist macro required by pandoc >= 1.14
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% title / author / date
+\title{Re-envisioning a future in scholarly communication}
+\author{Chris H.J. Hartgerink}
+\date{2017-05-23}
+
+\begin{document}
+
+\maketitle
+
+\textbf{For the 2017 IFLA conference, theme: Being open about open}
+
+\textbf{Made available under a
+\href{https://creativecommons.org/publicdomain/zero/1.0/legalcode}{CC 0
+public domain dedication}}
+
+\section{Abstract}\label{abstract}
+
+Scholarly communication is in need of disruption. Commodifying knowledge
+as is currently done with journals, is not sustainable any longer. An
+alternative is the commodification of how information is consumed. By
+focusing on the commodification of consumption instead of
+commodification of the resource, the problem of access to knowledge can
+be resolved in a sustainable manner. Additionally, commodification of
+consumption removes several perverse incentives from the scholarly
+system that now produces unreliable knowledge. The main tenet underlying
+the themes of Open Access, Open Data, Open Science, and replication
+initiatives in scholarly communication is sustainability through
+transparency of the scholarly process in all facets. The sustainability
+of any networked system is threatened by single points of failure (i.e.,
+the entire system can be manipulated from one node in the network). The
+scholarly process is ridden with such single points of failures at all
+stages. Distributing the scholarly communications system would remove
+the problems of single points of failure. Distributing and
+decentralizing the scholarly communications system is achievable with
+newly developed peer-to-peer (p2p) Internet protocols. Alongside
+decentralization and distribution of the content, integrity of the
+scholarly record can also be reformed to transform sections of a paper
+into different, reusable nodes of knowledge. These nodes can be logged
+on a blockchain based ledger of which everyone can have a copy. In order
+to deposit nodes onto the ledger, the depositor needs to agree that the
+contents are licensed CC 0, in order to maximize legal certainty
+regarding reuse of the contents. This is key to create a sustainable
+eco-system where scholars and companies can cooperate instead of
+compete, as we currently do.
+
+\section{Body}\label{body}
+
+Scholarly communication is in (dire) need of disruption. Subscription
+costs to scholarly journals are untenable already
+\citep[e.g.,][]{harvard-serials-crisis} and will become even more
+untenable in the long run with above inflation price-increases
+\citep[i.e., serials crisis;][]{serials-crisis}. The current scholarly
+incentive system detracts validity instead of adding it
+\citep[e.g.,][]{10.1371/journal.pmed.0020124, 10.1371/journal.pmed.0050201}
+by promoting behavior that is antithetical to the norms of science
+\citep{merton1942, 10.2307/2094423, 10.1525/jer.2007.2.4.3}. For
+example, results particular to one researcher instead of universal among
+researchers, because results cannot systematically be verified
+considering data are rarely shared
+\citep{10.1038/nature.2017.21549, 10.1525/collabra.13, 10.1037/0003-066x.61.7.726}
+or preserved \citep{10.1016/j.cub.2013.11.014}. Widely publicized
+results regularly fail reproducibility tests
+\citetext{\citealp[e.g.,][]{10.1126/science.aac4716}; \citealp{10.1038/483531a}; \citealp{10.1038/541269a}; \citealp[but
+see also][]{10.1016/j.jesp.2015.10.001}}, in part because published
+results are highly pre-selected for unscientific reasons
+\citetext{\citealp[e.g., pure novelty of
+results,][]{10.1016/j.tree.2013.05.007}; \citealp[easy to assimilate
+findings,][]{10.1038/526182a}; \citealp[statistical
+significance,][]{10.1126/science.1255484}; \citealp{10.1016/0140-67369190201-Y}}.
+These problems are not new but they are also not absolute --- scholars
+want to behave in accordance with the scientific norms
+\citep{10.1525/jer.2007.2.4.3} but the system seems to encourage us to
+do the opposite.
+
+In this perspective piece, I try to radically reimagine the scholarly
+communications infrastructure based on the digital tools that are
+available to us now, instead of the analog era and its legacies. I
+thoroughly believe that the Mertonian norms in science
+\citep{merton1942} align perfectly with transparency in science
+\citep[see also][]{10.14293/s2199-1006.1.sor-socsci.arysbi.v1} and that
+a scholarly communications system can be built on this framework in a
+sustainable manner. I invite everyone to criticize these ideas, for only
+with diversity of ideas can we truly make progress. Currently, how
+scholars operate (i.e., governance) is driven by a few oligopolies
+\citep{10.1371/journal.pone.0127502} or the highly pre-selected sample
+of tenured professors and policymakers, who mostly stem from the
+generation that has created this broken system and have benefited or
+still benefit from the current system. As a result, they are likely to
+underestimate the need for change in scholarly communications and cannot
+be expected to intervene in the necessary ways.
+
+In the last decades, interpersonal communication has changed
+tremendously. Not only has it brought much diversity in communication,
+so much diversity has come into existence that it has become confusing
+(see Figure 1). At the same time, that same diversity has liberated us
+by allowing innovative ways to communicate (e.g., social media) or
+making available previously difficult to use safety measures easily
+accessible to the masses (e.g., encryption). Diversity in
+communication-channels has thus spawned new ways of communicating.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.6\linewidth]{figs/xkcd_1810-cc-by-nc-2_5}
+
+}
+
+\caption[_Figure 1._ A comical example of the diversity in communication channels]{_Figure 1._ A comical example of the diversity in communication channels.
+[Licensed under CC BY-NC 2.5 from xkcd](https://xkcd.com/1810/)}\label{fig:xkcd-1810}
+\end{figure}
+
+The current scholarly communications system requires submitting to a
+journal that fits all scholarly output into a static, page-by-page, and
+paper format without the innovative solutions that are now available.
+This bottleneck not only prevents new ways of communicating findings, it
+also prevents further encouragement of discussion amongst readers about
+the findings. As a result, the current system encourages scholars to
+passively consume findings instead of critically evaluating them.
 
 Hence, we are stuck in a paper paradigm, albeit writing digital pages.
-Such digital pages radically restrict the way we imagine the communication of our results.
-In recent years, options to break free from this paper paradigm have arisen with tools such as [Rmarkdown](http://rmarkdown.rstudio.com/) or [Jupyter notebooks](https://jupyter.org/).
-These options allow the user to include code to dynamically generate figures, tables, or values in the text.
-By extension, how data are cleaned and analyzed can also be included as part of a manuscript for co-authors to verify or readers to discuss [see also @elife-rmd].
-Moreover, interactive elements can be embedded, such as interactive figures, online applications, or videos.
+Such digital pages radically restrict the way we imagine the
+communication of our results. In recent years, options to break free
+from this paper paradigm have arisen with tools such as
+\href{http://rmarkdown.rstudio.com/}{Rmarkdown} or
+\href{https://jupyter.org/}{Jupyter notebooks}. These options allow the
+user to include code to dynamically generate figures, tables, or values
+in the text. By extension, how data are cleaned and analyzed can also be
+included as part of a manuscript for co-authors to verify or readers to
+discuss \citep[see also][]{elife-rmd}. Moreover, interactive elements
+can be embedded, such as interactive figures, online applications, or
+videos.
 
-The scholarly process is much more complicated than can be captured in a single static and retrospective publication.
-The previously mentioned interactive elements enrich a single publication, but multiple moments of communicating ideas and results maps the empirical cycle onto the communication process.
-For example, a quantitative scholar needs to develop methods to investigate a research question.
-Feedback from peers would be valuable at that point.
-With so called Registered Reports [@10.1016/j.cortex.2012.12.016] a two-staged submission process is introduced to move away from a single static publication.
-It allows selected peers (i.e., peer reviewers) to discuss the contents at an interim stage.
-Additionally, this ensures chronology of predictions such as pre-registrations aim to do [@10.1038/s41562-016-0021], but in a more natural way.
-Nonetheless, registered reports, pre-registration, and data sharing are for the proactive, hence, remain the exception rather than the norm [@10.1371/journal.pbio.1002456], while these are already implicitly part of how scholarly research is conducted.
-Additionally, researchers create materials, protocols, and other outputs during the research process, share them with colleagues/advisors, but this process is not (aptly) sharable in a static publication.
-If the scholarly research process has more moments of communication, these implicit parts are finally able to show themselves.
+The scholarly process is much more complicated than can be captured in a
+single static and retrospective publication. The previously mentioned
+interactive elements enrich a single publication, but multiple moments
+of communicating ideas and results maps the empirical cycle onto the
+communication process. For example, a quantitative scholar needs to
+develop methods to investigate a research question. Feedback from peers
+would be valuable at that point. With so called Registered Reports
+\citep{10.1016/j.cortex.2012.12.016} a two-staged submission process is
+introduced to move away from a single static publication. It allows
+selected peers (i.e., peer reviewers) to discuss the contents at an
+interim stage. Additionally, this ensures chronology of predictions such
+as pre-registrations aim to do \citep{10.1038/s41562-016-0021}, but in a
+more natural way. Nonetheless, registered reports, pre-registration, and
+data sharing are for the proactive, hence, remain the exception rather
+than the norm \citep{10.1371/journal.pbio.1002456}, while these are
+already implicitly part of how scholarly research is conducted.
+Additionally, researchers create materials, protocols, and other outputs
+during the research process, share them with colleagues/advisors, but
+this process is not (aptly) sharable in a static publication. If the
+scholarly research process has more moments of communication, these
+implicit parts are finally able to show themselves.
 
-Additionally, the scholarly process has grown more complicated with the rise of research tools available and the speed with which alternative results can be inspected.
-Point-and-click software has increased the number of statistical analyses a researcher can do within minutes.
-This increased researcher degrees of freedom to present results that are more in line with previous results, leading to confirmation bias or motivated reasoning [@10.1037/0033-2909.108.3.480], hence, invalid results [@10.1177/0956797611417632].
-Moreover, the development of technology has created new types of data output that greatly exceed human comprehension and require algorithms before being able to make sense of the data (e.g., fMRI data).
-How a researcher applies all these tools is essential, but the current scholarly communication system does not require communication of the details needed to understand what steps were taken.
-Lack of direct reproducibility of results [e.g., @10.1101/056473] therefore is no surprise when we do not know how data are processed and analyzed.
-As such, the scholarly process has become more of a black box than it was before because communication has not adapted proportionate to the increasing complexity of the message that needs to be communicated.
-Increasing complexity and lack of transparency makes it increasingly difficult to reproduce steps taken, simply due to an increasing garden of forking paths [@10.1511/2014.111.460].
+Additionally, the scholarly process has grown more complicated with the
+rise of research tools available and the speed with which alternative
+results can be inspected. Point-and-click software has increased the
+number of statistical analyses a researcher can do within minutes. This
+increased researcher degrees of freedom to present results that are more
+in line with previous results, leading to confirmation bias or motivated
+reasoning \citep{10.1037/0033-2909.108.3.480}, hence, invalid results
+\citep{10.1177/0956797611417632}. Moreover, the development of
+technology has created new types of data output that greatly exceed
+human comprehension and require algorithms before being able to make
+sense of the data (e.g., fMRI data). How a researcher applies all these
+tools is essential, but the current scholarly communication system does
+not require communication of the details needed to understand what steps
+were taken. Lack of direct reproducibility of results
+\citep[e.g.,][]{10.1101/056473} therefore is no surprise when we do not
+know how data are processed and analyzed. As such, the scholarly process
+has become more of a black box than it was before because communication
+has not adapted proportionate to the increasing complexity of the
+message that needs to be communicated. Increasing complexity and lack of
+transparency makes it increasingly difficult to reproduce steps taken,
+simply due to an increasing garden of forking paths
+\citep{10.1511/2014.111.460}.
 
-<!-- Moreover, much of the software is proprietary and it is simply impossible to verify results or understand how they originated [@10.1038/nature10836].
-Trying to reverse engineer these pieces of software is actually often illegal based on the terms of service agreement that you have to accept before using the software.
-As such, it is practically impossible to understand how software development affects how results are computed, while testing for this is easy with open-source software.
--->
+Due to the fact that these complexities currently cannot and therefore
+are not reported, breaking the reign of the paper paradigm is a
+necessity to better represent and understand how scholarly results come
+into existence. Static paper publication remains a valid communication
+medium and it is not made obsolete when scholarly communications become
+truly digitally born and continuous. PDF versions of outputs can still
+be produced if a paper format is suitable or desirable. However, it is
+easy to imagine videos being embedded for protocols to be easily
+clarified (having ``Video X'' alongside ``Figure X'' and ``Table X'',
+for example) by origin, to then be replaced by stills when a paper
+version is produced. Nonetheless, without shifting the main focus from
+paper-based to digital-based scholarly communication, we cannot break
+free from the legacies of the paper paradigm. Paper-based communication
+includes the legacy of printing research outputs by default, which by
+extension includes the idea of subscriptions, purchasing content, and
+restricting the flow of information. Copyright is the legacy of the
+paper-based age that seems to limit the digital potential of scholarly
+outputs. As such, the legacy of a paper-based paradigm with respect to
+scholarly communication is that of knowledge commodification.
 
-Due to the fact that these complexities currently cannot and therefore are not reported, breaking the reign of the paper paradigm is a necessity to better represent and understand how scholarly results come into existence.
-Static paper publication remains a valid communication medium and it is not made obsolete when scholarly communications become truly digitally born and continuous.
-PDF versions of outputs can still be produced if a paper format is suitable or desirable.
-However, it is easy to imagine videos being embedded for protocols to be easily clarified (having "Video X" alongside "Figure X" and "Table X", for example) by origin, to then be replaced by stills when a paper version is produced.
-Nonetheless, without shifting the main focus from paper-based to digital-based scholarly communication, we cannot break free from the legacies of the paper paradigm.
-Paper-based communication includes the legacy of printing research outputs by default, which by extension includes the idea of subscriptions, purchasing content, and restricting the flow of information.
-Copyright is the legacy of the paper-based age that seems to limit the digital potential of scholarly outputs.
-As such, the legacy of a paper-based paradigm with respect to scholarly communication is that of knowledge commodification.
+For scholarly communication, commodifying knowledge as it is currently
+done can be seen as a human rights violation.
+\href{http://wayback.archive.org/web/20170510094028/http://www.un.org/en/universal-declaration-human-rights/index.html}{The
+Universal Declaration of Human Rights (UDHR), Article 27 section 1},
+reads as follows
 
-For scholarly communication, commodifying knowledge as it is currently done can be seen as a human rights violation.
-[The Universal Declaration of Human Rights (UDHR), Article 27 section 1](http://wayback.archive.org/web/20170510094028/http://www.un.org/en/universal-declaration-human-rights/index.html), reads as follows
+\begin{quote}
+Everyone has the right freely to participate in the cultural life of the
+community, to enjoy the arts and to share in scientific advancement and
+its benefits.
+\end{quote}
 
-> Everyone has the right freely to participate in the cultural life of the community, to enjoy the arts and to share in scientific advancement and its benefits.
+When knowledge is commodified, it results in decreased participation,
+enjoyment, and sharing of scholarly progress. It could therefore be
+argued that this is a violation of human rights \citep{udhr}. Moreover,
+this is the main legal motivation proposed by Alexandra Elbakyan as to
+why \href{http://sci-hub.bz}{Sci-Hub} is not illegal from her point of
+view \citep{alexandraelbakyan}. Discussion on the legal correctness of
+this argument would be worthwhile to further learn about this. On the
+other hand, legal cases on this could result in clarification of whether
+unauthorized sharing, which is part of scholarly culture
+\citep{10.1038/545145a}, is allowed from a human rights perspective and
+whether such sharing would supercede copyright legislation.
 
-When knowledge is commodified, it results in decreased participation, enjoyment, and sharing of scholarly progress.
-It could therefore be argued that this is a violation of human rights [@udhr].
-Moreover, this is the main legal motivation proposed by Alexandra Elbakyan as to why [Sci-Hub](http://sci-hub.bz) is not illegal from her point of view [@alexandraelbakyan].
-Discussion on the legal correctness of this argument would be worthwhile to further learn about this.
-On the other hand, legal cases on this could result in clarification of whether unauthorized sharing, which is part of scholarly culture [@10.1038/545145a], is allowed from a human rights perspective and whether such sharing would supercede copyright legislation.
+Moreover, commodifying knowledge is not the only way to make scholarly
+communication sustainable; an alternative is the commodification of how
+information is consumed. That is, commodifying the services that serve
+the users the content, instead of the content itself. For example, if
+all scholarly information is freely available, (not-)for-profit
+organizations can compete for users who want to sift through that
+information efficiently. If an algorithm is developed that greatly
+increases efficiency of finding relevant articles, revenue could be
+built with usage fees to that service instead of the content. Such a
+market could create competition in information consumption, which is
+timely given the information explosion humanity is going through at the
+moment. The potential market capitalization of free flow of information
+and data is tremendous \citep[e.g., freely reusable public sector
+information/data in the U.K. is estimated to provide a market cap of
+£590 million - £16 billion; p.~96 of][]{dotecon}. A freely reusable
+information resource would provide the infrastructure that encourages
+innovation and competition as to how people can discover and consume
+that information.
 
-Moreover, commodifying knowledge is not the only way to make scholarly communication sustainable; an alternative is the commodification of how information is consumed.
-That is, commodifying the services that serve the users the content, instead of the content itself.
-For example, if all scholarly information is freely available, (not-)for-profit organizations can compete for users who want to sift through that information efficiently.
-If an algorithm is developed that greatly increases efficiency of finding relevant articles, revenue could be built with usage fees to that service instead of the content.
-Such a market could create competition in information consumption, which is timely given the information explosion humanity is going through at the moment.
-The potential market capitalization of free flow of information and data is tremendous [e.g., freely reusable public sector information/data in the U.K.
-is estimated to provide a market cap of £590 million - £16 billion; p.
-96 of @dotecon].
-A freely reusable information resource would provide the infrastructure that encourages innovation and competition as to how people can discover and consume that information.
+Consequently, by focusing on the commodification of consumption instead
+of commodification of the resource, the problem of access to knowledge
+can be resolved in a sustainable manner. When the consumption is
+commodified, all service providers benefit from creating the largest
+pool of resources that is freely available to build their service on. By
+extension, it shifts the optimal outcome from not sharing (i.e.,
+defecting) to sharing (i.e., cooperating), because it increases the
+value of their service. This would create a market that stimulates
+access instead of limiting it, while retaining the potential for
+revenue.
 
-Consequently, by focusing on the commodification of consumption instead of commodification of the resource, the problem of access to knowledge can be resolved in a sustainable manner.
-When the consumption is commodified, all service providers benefit from creating the largest pool of resources that is freely available to build their service on.
-By extension, it shifts the optimal outcome from not sharing (i.e., defecting) to sharing (i.e., cooperating), because it increases the value of their service.
-This would create a market that stimulates access instead of limiting it, while retaining the potential for revenue.
+Additionally, commodification of consumption removes several perverse
+incentives from the scholarly system that now produces unreliable
+knowledge. For example, innovativeness and surprisingness are criteria
+upon which articles are currently selected, where those exact some
+properties increase the probability that the finding is false
+\citep{10.1371/journal.pmed.0020124}. When the knowledge is commodified,
+this makes sense from a business perspective (comparable to news
+outlets): readers will buy access more to shocking and astounding
+results. However, when consumption is commodified, this whole incentive
+is removed and the most valuable resource is the most populated and
+diverse one. This would allow much more information to be extracted,
+hence, richer services to be built on top of that resource.
 
-Additionally, commodification of consumption removes several perverse incentives from the scholarly system that now produces unreliable knowledge.
-For example, innovativeness and surprisingness are criteria upon which articles are currently selected, where those exact some properties increase the probability that the finding is false [@10.1371/journal.pmed.0020124].
-When the knowledge is commodified, this makes sense from a business perspective (comparable to news outlets): readers will buy access more to shocking and astounding results.
-However, when consumption is commodified, this whole incentive is removed and the most valuable resource is the most populated and diverse one.
-This would allow much more information to be extracted, hence, richer services to be built on top of that resource.
+The main tenet underlying the themes of Open Access, Open Data, Open
+Science, and replication initiatives in scholarly communication is
+sustainability through transparency of the scholarly process in all
+facets. As I tried to explain in the previous paragraphs, this
+sustainability can be achieved by recalibrating the fundamental business
+model of the scholarly communications system. In the next section of
+this perspective piece, I propose a redesign of the scholarly
+communication system that is flexible, decentralized, distributed, and
+freely accessible and reusable from the start. Such a system would allow
+for more sustainability and diversity in the way we consume, produce,
+and access knowledge, which would ultimately benefit everyone who reaps
+the benefits of scholarly results.
 
-The main tenet underlying the themes of Open Access, Open Data, Open Science, and replication initiatives in scholarly communication is sustainability through transparency of the scholarly process in all facets.
-As I tried to explain in the previous paragraphs, this sustainability can be achieved by recalibrating the fundamental business model of the scholarly communications system.
-In the next section of this perspective piece, I propose a redesign of the scholarly communication system that is flexible, decentralized, distributed, and freely accessible and reusable from the start.
-Such a system would allow for more sustainability and diversity in the way we consume, produce, and access knowledge, which would ultimately benefit everyone who reaps the benefits of scholarly results.
+\section{Sustainable scholarly
+communication}\label{sustainable-scholarly-communication}
 
-## Sustainable scholarly communication
+The sustainability of any networked system is threatened by single
+points of failure (i.e., the entire system can be manipulated from one
+node in the network). Single points of failure not only make a system
+vulnerable to downtime, it also raises the possibility of (malicious)
+adjustment of content, or even complete removal of content (among other
+things). For example, retracted papers sometimes wholly disappear,
+leaving only the retraction notice; this results in a direct and nearly
+irreversable adjustment of the scholarly record (from the scholar's
+perspective). Due to these single (or few) points of failure, such a
+centralized and concentrated infrastructure is easier to disrupt than a
+decentralized or distributed one \citep[see
+also][]{10.1103/physreve.95.022313}.
 
-The sustainability of any networked system is threatened by single points of failure (i.e., the entire system can be manipulated from one node in the network).
-Single points of failure not only make a system vulnerable to downtime, it also raises the possibility of (malicious) adjustment of content, or even complete removal of content (among other things).
-For example, retracted papers sometimes wholly disappear, leaving only the retraction notice; this results in a direct and nearly irreversable adjustment of the scholarly record (from the scholar's perspective).
-Due to these single (or few) points of failure, such a centralized and concentrated infrastructure is easier to disrupt than a decentralized or distributed one [see also @10.1103/physreve.95.022313].
+The scholarly process is ridden with single- or few points of failures
+at all stages. Hence its sustainability is directly threatened. As
+mentioned before, communication occurs only once (i.e., at the end);
+this could be regarded as a single point of failure for communicating
+what the scholars did. If data are not shared or preserved, the scholar
+creates a single point of failure in accessing and preserving the data.
+When results are published with a closed access publisher, the publisher
+becomes the single point of failure for providing access to those
+results. For exactly this reason I find it important that scholars
+become more transparent about their data, their hypotheses, their
+results, and the chronology of their research process; it decreases the
+risk of a single point of failure actually failing.
 
-The scholarly process is ridden with single- or few points of failures at all stages.
-Hence its sustainability is directly threatened.
-As mentioned before, communication occurs only once (i.e., at the end); this could be regarded as a single point of failure for communicating what the scholars did.
-If data are not shared or preserved, the scholar creates a single point of failure in accessing and preserving the data.
-When results are published with a closed access publisher, the publisher becomes the single point of failure for providing access to those results.
-For exactly this reason I find it important that scholars become more transparent about their data, their hypotheses, their results, and the chronology of their research process; it decreases the risk of a single point of failure actually failing.
+Distributing the scholarly communications system would remove the
+problems of single- or few points of failure in access to findings.
+Currently, \href{http://www.portico.org/digital-preservation/}{Portico}
+and \href{https://clockss.org/clockss/Home}{CLOCKSS} are supposed to
+provide access in case the original content is subject to a
+\href{https://clockss.org/clockss/FAQ\#How_does_the_CLOCKSS_board_define_a_trigger_event.3F}{``trigger
+event''} (e.g., publisher goes out of business). However, this system is
+insufficient for wide distribution because content is only selectively
+distributed to libraries. Moreover, by its nature ``trigger events'' are
+reactive and need governance (indicated by the Centralized LOCKSS
+system, i.e., CLOCKSS). That such governance is problematic, is
+subscribed by the case that CLOCKSS has previously preserved content
+under a more restrictive license than was originally used
+(\href{http://wayback.archive.org/web/20170518110938/https://clockss.org/clockss/Tijdschrift_voor_Tijdschriftstudies}{CC
+BY-NC-ND for the preserved journal ``Tijdschrift voor
+Tijdschriftstudies''}, whereas
+\href{http://triggered.stanford.clockss.org/ServeContent?url=https\%3A\%2F\%2Fwww.tijdschriftstudies.nl\%2Farticles\%2F10.18352\%2Fts.340\%2Fprint\%2F}{the
+original was CC BY}). In order to make a truly distributed system
+sustainable, persistent nodes in the network need to be identified to
+store all this information (e.g., libraries), but institutions and
+individuals should be able to freely participate in the distributed
+network by freely creating temporary or persistent copies for wide and
+unlimited distribution of content.
 
-Distributing the scholarly communications system would remove the problems of single- or few points of failure in access to findings.
-Currently, [Portico](http://www.portico.org/digital-preservation/) and [CLOCKSS](https://clockss.org/clockss/Home) are supposed to provide access in case the original content is subject to a ["trigger event"](https://clockss.org/clockss/FAQ#How_does_the_CLOCKSS_board_define_a_trigger_event.3F) (e.g., publisher goes out of business).
-However, this system is insufficient for wide distribution because content is only selectively distributed to libraries.
-Moreover, by its nature "trigger events" are reactive and need governance (indicated by the Centralized LOCKSS system, i.e., CLOCKSS).
-That such governance is problematic, is subscribed by the case that CLOCKSS has previously preserved content under a more restrictive license than was originally used ([CC BY-NC-ND for the preserved journal "Tijdschrift voor Tijdschriftstudies"](http://wayback.archive.org/web/20170518110938/https://clockss.org/clockss/Tijdschrift_voor_Tijdschriftstudies), whereas [the original was CC BY](http://triggered.stanford.clockss.org/ServeContent?url=https%3A%2F%2Fwww.tijdschriftstudies.nl%2Farticles%2F10.18352%2Fts.340%2Fprint%2F)).
-In order to make a truly distributed system sustainable, persistent nodes in the network need to be identified to store all this information (e.g., libraries), but institutions and individuals should be able to freely participate in the distributed network by freely creating temporary or persistent copies for wide and unlimited distribution of content.
+Distributing and decentralizing the scholarly communications system is
+achievable with peer-to-peer (p2p) Internet protocols such as
+\href{https://datproject.org/}{\texttt{dat}} and
+\href{https://ipfs.io/}{\texttt{ipfs}}. Simply put, such p2p networks
+securely send information across a network of peers but are resilient to
+nodes being removed or adjusted because they operate in a mesh network.
+For example, if 20 peers have file X, removing one peer does not affect
+the availability of the file X. Only if all 20 are removed from the
+network, file X will become unavailable. Vice versa, if more peers on
+the network have file X, it is less likely that file X will become
+unavailable. As such, this would include unlimited redistribution in the
+scholarly communication system by default, instead of limited
+redistribution due to copyright as it is now.
 
-Distributing and decentralizing the scholarly communications system is achievable with peer-to-peer (p2p) Internet protocols such as [`dat`](https://datproject.org/) and [`ipfs`](https://ipfs.io/).
-Simply put, such p2p networks securely send information across a network of peers but are resilient to nodes being removed or adjusted because they operate in a mesh network.
-For example, if 20 peers have file X, removing one peer does not affect the availability of the file X.
-Only if all 20 are removed from the network, file X will become unavailable.
-Vice versa, if more peers on the network have file X, it is less likely that file X will become unavailable.
-As such, this would include unlimited redistribution in the scholarly communication system by default, instead of limited redistribution due to copyright as it is now.
+Alongside decentralization and distribution of the content as described
+in the previous paragraphs, integrity of the scholarly record can also
+be reformed to transform sections of a paper into different, reusable
+nodes of knowledge. The chronology of the scholarly process can be
+ensured by splitting up the paper article as we know it into nodes of
+information. The theory section could be a node, from which a hypotheses
+node follows, from which a methods node follows, and so on. This would
+provide a more true depiction of the scholarly process and facilitate
+direct replications (one needs to only take the materials node and
+create a new data node), reanalyses (one needs to only take the data
+node and create a new results node), etc. Additionally, scholars would
+be able to start citing more granularly by referring to specific nodes
+instead of entire papers and would not have to write the same theory
+section (for example) multiple times for different papers.
 
-Alongside decentralization and distribution of the content as described in the previous paragraphs, integrity of the scholarly record can also be reformed to transform sections of a paper into different, reusable nodes of knowledge.
-The chronology of the scholarly process can be ensured by splitting up the paper article as we know it into nodes of information.
-The theory section could be a node, from which a hypotheses node follows, from which a methods node follows, and so on.
-This would provide a more true depiction of the scholarly process and facilitate direct replications (one needs to only take the materials node and create a new data node), reanalyses (one needs to only take the data node and create a new results node), etc.
-Additionally, scholars would be able to start citing more granularly by referring to specific nodes instead of entire papers and would not have to write the same theory section (for example) multiple times for different papers.
+These nodes can be logged on a blockchain based ledger (e.g.,
+\href{https://github.com/hyperledger/fabric}{Hyperledger}-based) of
+which everyone can have a copy. Although explaining the entire operating
+principle of the blockchain is beyond the scope of this piece, the
+essence is relatively simple. Each entry on the blockchain needs to
+satisfy a predefined mathematical rule, which is dependent on the
+contents of the previous entry in the ledger, the contents of the new
+entry, and a constant to offset the current entry \citep{bitcoin}.
+Creating a situation that fullfills that mathematical rule requires
+solving for the offset in the new entry to the blockchain, which is
+called `mining'. Because the proof of record (i.e., satisfying the
+predefined mathematical rule) is dependent on the combination between
+the contents of the previous- and current entries (i.e., blocks) and the
+offset, the chain is broken when the contents of any entry in that chain
+are (unwillingly) changed. Given that each user has a copy of the entire
+blockchain, those who have a different chain from the majority of other
+users are simply rejected. As such, in order to change the ledger, each
+change would have to be re-verified by the majority of the network
+(i.e., \textgreater{}50\%) to become verified (for a good introduction
+to the blockchain, see
+\url{https://www.youtube.com/watch?v=_160oMzblY8}).
 
-These nodes can be logged on a blockchain based ledger (e.g., [Hyperledger](https://github.com/hyperledger/fabric)-based) of which everyone can have a copy.
-Although explaining the entire operating principle of the blockchain is beyond the scope of this piece, the essence is relatively simple.
-Each entry on the blockchain needs to satisfy a predefined mathematical rule, which is dependent on the contents of the previous entry in the ledger, the contents of the new entry, and a constant to offset the current entry [@bitcoin].
-Creating a situation that fullfills that mathematical rule requires solving for the offset in the new entry to the blockchain, which is called 'mining'.
-Because the proof of record (i.e., satisfying the predefined mathematical rule) is dependent on the combination between the contents of the previous- and current entries (i.e., blocks) and the offset, the chain is broken when the contents of any entry in that chain are (unwillingly) changed.
-Given that each user has a copy of the entire blockchain, those who have a different chain from the majority of other users are simply rejected.
-As such, in order to change the ledger, each change would have to be re-verified by the majority of the network (i.e., >50%) to become verified (for a good introduction to the blockchain, see [https://www.youtube.com/watch?v=_160oMzblY8](https://www.youtube.com/watch?v=_160oMzblY8)).
+Figure 2 shows an example of how the nodes might be incorporated into
+such a blockchain based ledger. The idea behind this format is that it
+needs to contain the most crucial information, but also be flexible to
+accompany different demands in different scholarly fields and
+development in scholarly research over time. The ``parent'' entry
+indicates what the related nodes are, which chronologically predate the
+new node. This is not meant to replace citations, but only to indicate
+direct links between nodes if applicable. For example, there might be a
+``data'' node that spawns competing research teams to analyze the data
+in different ways, creating two ``results'' nodes that come from the
+same ``data'' node. Subsequently, each node can itself become a parent
+at a later point, which is why the ``self'' entry is included. The
+``parent'' and ``self'' can be any p2p address from \texttt{dat},
+\texttt{ipfs} or future p2p platform (i.e., this approach is platform
+agnostic). This efficiently combines the reference to the previous nodes
+and its contents, and the contents of the new node.
 
-Figure 2 shows an example of how the nodes might be incorporated into such a blockchain based ledger.
-The idea behind this format is that it needs to contain the most crucial information, but also be flexible to accompany different demands in different scholarly fields and development in scholarly research over time.
-The "parent" entry indicates what the related nodes are, which chronologically predate the new node.
-This is not meant to replace citations, but only to indicate direct links between nodes if applicable.
-For example, there might be a "data" node that spawns competing research teams to analyze the data in different ways, creating two "results" nodes that come from the same "data" node.
-Subsequently, each node can itself become a parent at a later point, which is why the "self" entry is included.
-The "parent" and "self" can be any p2p address from `dat`, `ipfs` or future p2p platform (i.e., this approach is platform agnostic).
-This efficiently combines the reference to the previous nodes and its contents, and the contents of the new node.
+\begin{figure}
+\includegraphics[width=1\linewidth]{figs/fig-json} \caption[_Figure 2._ An example of how two nodes might be described and stored in the blockchain and a depiction of how these two entries relate]{_Figure 2._ An example of how two nodes might be described and stored in the blockchain and a depiction of how these two entries relate.}\label{fig:fig_json}
+\end{figure}
 
-```{r fig_json, fig.cap="_Figure 2._ An example of how two nodes might be described and stored in the blockchain and a depiction of how these two entries relate.", echo=FALSE, out.width="100%"}
-knitr::include_graphics('figs/fig-json.png')
-```
+The nodes also include ``time-deposit'' and ``time-available'' entries;
+these indicate the time-related aspects of the node. To ensure
+chronology, the time the node is deposited into the ledger is included,
+ensuring that if used as a parent, it precedes the child node.
+Additionally, it makes sense that not all information can be made
+available at the same time as the deposit takes place, which is why
+``time-available'' is included to make clear when that information will
+become available. There are various ways in which this time-locked
+system could be implemented. Either the author can bear the
+responsibility to deposit the materials in a timely manner, or a trusted
+third-party can provide a service to automatically deposit these
+materials. Alternatively, files could be encrypted and deposited at the
+same time as the entry to the ledger is made, where the private key
+needed to decrypt the files could be provided at the ``time-available''.
+A system that incorporates ``time-available'' moves the discussion of
+whether materials should be shared at all to when they can be shared
+(i.e., not much can be justified to stay unavailable into perpetuity).
 
-The nodes also include "time-deposit" and "time-available" entries; these indicate the time-related aspects of the node.
-To ensure chronology, the time the node is deposited into the ledger is included, ensuring that if used as a parent, it precedes the child node.
-Additionally, it makes sense that not all information can be made available at the same time as the deposit takes place, which is why "time-available" is included to make clear when that information will become available.
-There are various ways in which this time-locked system could be implemented.
-Either the author can bear the responsibility to deposit the materials in a timely manner, or a trusted third-party can provide a service to automatically deposit these materials.
-Alternatively, files could be encrypted and deposited at the same time as the entry to the ledger is made, where the private key needed to decrypt the files could be provided at the "time-available".
-A system that incorporates "time-available" moves the discussion of whether materials should be shared at all to when they can be shared (i.e., not much can be justified to stay unavailable into perpetuity).
-
-In order to provide identification and accountability of nodes, author ORCIDs can be combined into unique author combinations, which also diminishes gender and status bias during initial assessment (amongst other things).
-These combinations can be created based on so-called hashes (e.g., SHA256SUM with salting).
-For example, my ORCID is [`0000-0003-1050-6809`](http://orcid.org/0000-0003-1050-6809) and that of one of my supervisors is [`0000-0003-2415-2933`](http://orcid.org/0000-0003-2415-2933).
-If I am the sole author, my "authors" hash would be ``r digest::digest('0000-0003-1050-6809', 'sha256')``, but if I am first author and my supervisor is second author, the hash is different (i.e., ``r digest::digest('0000-0003-1050-6809;0000-0003-2415-2933', 'sha256')``).
-If we do not disclose these hashes or that we authored these nodes in some composition, it is practically impossible to reverse the hash into the identifiable ORCIDs.
-These hashes are not meant as a foolproof way to safeguard author identity and remove biases in an absolute manner; these hashes are primarily meant as a way to reduce the size of each entry in the ledger (i.e., the hash is of the same length regardless of how many authors are included) and to obfuscate the author's identity.
+In order to provide identification and accountability of nodes, author
+ORCIDs can be combined into unique author combinations, which also
+diminishes gender and status bias during initial assessment (amongst
+other things). These combinations can be created based on so-called
+hashes (e.g., SHA256SUM with salting). For example, my ORCID is
+\href{http://orcid.org/0000-0003-1050-6809}{\texttt{0000-0003-1050-6809}}
+and that of one of my supervisors is
+\href{http://orcid.org/0000-0003-2415-2933}{\texttt{0000-0003-2415-2933}}.
+If I am the sole author, my ``authors'' hash would be
+\texttt{a63d529d09b7c9de10e2e9cd71d51acec2f33f4f01af09eda5da152e0730f33f},
+but if I am first author and my supervisor is second author, the hash is
+different (i.e.,
+\texttt{494a82a472b301e95a08b9334fc962a03f7cb196c72c10d997e75cf0d5c00de7}).
+If we do not disclose these hashes or that we authored these nodes in
+some composition, it is practically impossible to reverse the hash into
+the identifiable ORCIDs. These hashes are not meant as a foolproof way
+to safeguard author identity and remove biases in an absolute manner;
+these hashes are primarily meant as a way to reduce the size of each
+entry in the ledger (i.e., the hash is of the same length regardless of
+how many authors are included) and to obfuscate the author's identity.
 Such obfuscation is helpful to reduce bias in peer assessments.
@@
 What types of nodes are contained in the ledger is up for discussion.
-For example, it is relatively clear that there can be "theory", "hypothesis", "methods", "data", "analysis", "results", and "discussion" nodes to cover the largest part of empirical research.
-However, there are various other types of nodes that will not fit into this rather restrictive taxonomy (e.g., "opinion" or those related to qualitative research).
-Moreover, over time, new types of nodes might be the result of natural development in scholarly research (e.g., due to additions or changes to the scientific method).
-As such, the system needs to be dynamic about the types of nodes it can include, but there should also not be total chaos in the types of nodes.
-This node taxonomy will require empirical study within the scholarly community to provide an evidence-base for an initial version.
+For example, it is relatively clear that there can be ``theory'',
+``hypothesis'', ``methods'', ``data'', ``analysis'', ``results'', and
+``discussion'' nodes to cover the largest part of empirical research.
+However, there are various other types of nodes that will not fit into
+this rather restrictive taxonomy (e.g., ``opinion'' or those related to
+qualitative research). Moreover, over time, new types of nodes might be
+the result of natural development in scholarly research (e.g., due to
+additions or changes to the scientific method). As such, the system
+needs to be dynamic about the types of nodes it can include, but there
+should also not be total chaos in the types of nodes. This node taxonomy
+will require empirical study within the scholarly community to provide
+an evidence-base for an initial version.
 
-The contents of each node can vary, depending on the type of node.
-For example, one node can refer to just one file directly (e.g., a `csv` dataset), whereas others can refer to entire folders, containing sets of files.
-For example, a data node can contain multiple data files if it refers to a folder, or a theory section can contain the text in various (open) formats, such as Markdown and XML.
-Moreover, nodes on p2p networks such as `dat` by default support large files (e.g., 50GB); such large file storage is becoming increasingly necessary with the large amount of data available.
+The contents of each node can vary, depending on the type of node. For
+example, one node can refer to just one file directly (e.g., a
+\texttt{csv} dataset), whereas others can refer to entire folders,
+containing sets of files. For example, a data node can contain multiple
+data files if it refers to a folder, or a theory section can contain the
+text in various (open) formats, such as Markdown and XML. Moreover,
+nodes on p2p networks such as \texttt{dat} by default support large
+files (e.g., 50GB); such large file storage is becoming increasingly
+necessary with the large amount of data available.
 
-In order to deposit nodes onto the ledger, the depositor needs to agree that the contents are licensed CC 0, in order to maximize legal certainty regarding reuse of the contents.
-The current copyright system has created the problem of inaccessibility and CC 0 empowers access by default.
-Moreover, restrictive licensing also comes with duties to enforce the license, if it is to mean anything to begin with.
-Enforcing a restrictive license requires scholars to legally pursue those who infringe their copyrights, detracting from their research time.
-As such, scholars are not the ones who benefit from restrictive licensing, and neither does the scholarly community at large.
+In order to deposit nodes onto the ledger, the depositor needs to agree
+that the contents are licensed CC 0, in order to maximize legal
+certainty regarding reuse of the contents. The current copyright system
+has created the problem of inaccessibility and CC 0 empowers access by
+default. Moreover, restrictive licensing also comes with duties to
+enforce the license, if it is to mean anything to begin with. Enforcing
+a restrictive license requires scholars to legally pursue those who
+infringe their copyrights, detracting from their research time. As such,
+scholars are not the ones who benefit from restrictive licensing, and
+neither does the scholarly community at large.
 
-# Conclusion
+\chapter{Conclusion}\label{conclusion}
 
-At OpenCon2016, Brewster Kahle mentioned that the scholarly system should be "locked open".
-After this talk I spent much time thinking about how this could be done.
-Using a decentralized and distributed system, as I tried to conceptualize throughout this article, is my initial attempt at realizing a "locked open" system that benefits not just the scholarly community, but also those that aim to generate value from these outputs.
-By shifting from a knowledge commodification to a commodification of how that knowledge is consumed, free access and reuse becomes beneficial to all parties.
-This is key to create a sustainable eco-system where scholars and companies can cooperate instead of compete, as we currently do.
+At OpenCon2016, Brewster Kahle mentioned that the scholarly system
+should be ``locked open''. After this talk I spent much time thinking
+about how this could be done. Using a decentralized and distributed
+system, as I tried to conceptualize throughout this article, is my
+initial attempt at realizing a ``locked open'' system that benefits not
+just the scholarly community, but also those that aim to generate value
+from these outputs. By shifting from a knowledge commodification to a
+commodification of how that knowledge is consumed, free access and reuse
+becomes beneficial to all parties. This is key to create a sustainable
+eco-system where scholars and companies can cooperate instead of
+compete, as we currently do.
 
-## References
+\section{References}\label{references}
+
+\bibliography{library.bib}
+
+\end{document}

### 2017-05-30 manuscript.Rmd
Update to submitted version

--- 2017-05-23 manuscript.Rmd
@@
       in_header: hypothesis.html
+  pdf_document: default
   word_document: default
@@
 By focusing on the commodification of consumption instead of commodification of the resource, the problem of access to knowledge can be resolved in a sustainable manner.
-Additionally, commodification of consumption removes several perverse incentives from the scholarly system that now produces unreliable knowledge.
+Additionally, commodification of consumption removes several perverse incentives from the scholarly system that now produce unreliable knowledge.
 The main tenet underlying the themes of Open Access, Open Data, Open Science, and replication initiatives in scholarly communication is sustainability through transparency of the scholarly process in all facets.
@@
 Additionally, researchers create materials, protocols, and other outputs during the research process, share them with colleagues/advisors, but this process is not (aptly) sharable in a static publication.
-If the scholarly research process has more moments of communication, these implicit parts are finally able to show themselves.
+If the scholarly research process has more moments of communication incorporating the many steps in research, these implicit parts are finally able to show themselves and allow credit to all types of researchers (e.g., theory builders, data-collectors, reanalyzers, etc.).
 
@@
 Moreover, the development of technology has created new types of data output that greatly exceed human comprehension and require algorithms before being able to make sense of the data (e.g., fMRI data).
-How a researcher applies all these tools is essential, but the current scholarly communication system does not require communication of the details needed to understand what steps were taken.
+How a researcher applies all these tools is essential, but the current scholarly communication system does not require communication of the details needed to understand what steps were taken, while at the same time neglecting the potential for education in (best) practices.
 Lack of direct reproducibility of results [e.g., @10.1101/056473] therefore is no surprise when we do not know how data are processed and analyzed.
@@
 Paper-based communication includes the legacy of printing research outputs by default, which by extension includes the idea of subscriptions, purchasing content, and restricting the flow of information.
-Copyright is the legacy of the paper-based age that seems to limit the digital potential of scholarly outputs.
+Using copyright to limit access to scholarly communication is the legacy of the paper-based age that limits the digital potential of scholarly outputs.
 As such, the legacy of a paper-based paradigm with respect to scholarly communication is that of knowledge commodification.
@@
 For example, if all scholarly information is freely available, (not-)for-profit organizations can compete for users who want to sift through that information efficiently.
-If an algorithm is developed that greatly increases efficiency of finding relevant articles, revenue could be built with usage fees to that service instead of the content.
-Such a market could create competition in information consumption, which is timely given the information explosion humanity is going through at the moment.
+If an algorithm is developed that greatly increases efficiency of finding relevant articles, this could generate revenue under various business models (e.g., subscription- or advertising model).
+Depending on the business model, this might shift inequality in access (current situation) to inequality in discovery.
+Nonetheless, a competitive market could create innovation in ways to consume information and puts pressure on the price levels to access these new services.
+New ways to consume information is timely given the information explosion humanity is going through at the moment.
 The potential market capitalization of free flow of information and data is tremendous [e.g., freely reusable public sector information/data in the U.K.
